### 2009-10-27 papers/2009-icon-hyderabad/paper.tex
Pridan zarodek clanku o indickem parsingu na ICON 2009 Hyderabad.

% !TEX TS-program = xelatex
% !TEX encoding = UTF-8 Unicode
% File acl-ijcnlp2009.tex
%
% Contact  jshin@csie.ncnu.edu.tw
%%
%% Based on the style files for EACL-2009 and IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl-ijcnlp2009}
\usepackage[	
   pdfdisplaydoctitle, breaklinks, colorlinks, linkcolor=black, citecolor=black, filecolor=black, urlcolor=black, 
   backref, hyperfootnotes]{hyperref} % backref a modre URL asi nakonec zrusime 
%\usepackage{times}
\usepackage{url}
\usepackage{amsmath}
\usepackage{color} %pro korektury
\usepackage{paralist} % for better itemize and enumerate

% a footer required for the first page
\usepackage{fancyhdr}
\fancyhead{} % clear all header fields
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[C]{Proceedings of ICON-2009: 7th International Conference on Natural Language Processing, Macmillan Publishers, India. Also accessible from http://ltrc.iiit.ac.in/proceedings/ICON-2009}

% xelatex
\usepackage{fontspec, xunicode, xltxtra}
\defaultfontfeatures{Mapping=tex-text}
\setmainfont{Times New Roman}
\setmonofont[Scale=MatchLowercase]{Luxi Mono}
\setmathsf{Lohit Hindi}%\XXX
\newfontinstance\hi[Script=Devanagari]{Lohit Hindi}
\newcommand{\hindi}[1]{{\hi #1}}
\newfontinstance\tl{Gentium}
\newcommand{\translit}[1]{{\tl \textit{#1}}}

% natbib
\usepackage{natbib}
\bibliographystyle{plainnat}
\bibpunct{(}{)}{;}{a}{,}{,}

% our defs
\def\perscite#1{\citet{#1}}  
\def\parcite#1{\citep{#1}} 
%ps: Did you mean \citep and \citet (in-parentheses and textual reference)? There is even more, see `texdoc natbib`.
%obo: your def seems right.
\def\Sref#1{Section~\ref{#1}}
\def\Tref#1{Table~\ref{#1}}
\def\Fref#1{Figure~\ref{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}} % komentare (TODO)
\newcommand{\XXX}{\textcolor{red}{XXX }} % komentare (TODO)

\title{English-Hindi Translation -- Obtaining Mediocre Results\\with Bad Data and
Fancy Models%
\thanks{ \hspace{.6em}The research has been supported by the grants GAAV ČR 1ET201120505 (Grant Agency of the Academy of Sciences of the Czech Republic), 
MSM0021620838 (Czech Ministry of Education),
FP7-ICT-2007-3-231720 (EuroMatrix Plus), and GAUK 4307/2009 (Grant agency of the Charles University).
}
}

% oboXXX nedat Gaurava az za Michaly?
% dz: jsem spíš proti. Konečný přínos Michalů je možná větší, ale pro Gaurava to byla ty 3 měsíce hlavní náplň činnosti a dělal přímo na samotném překladu, ne na "doplňcích".
\author{Ondřej Bojar, Pavel Straňák, Daniel Zeman, Gaurav Jain, Michal Hrušecký, Michal Richter, Jan Hajič\\
Univerzita Karlova v Praze, Ústav formální a aplikované lingvistiky,\\
Malostranské náměstí 25, 118 00, Praha, Czech Republic\\ 
\texttt{\{bojar,stranak,zeman,hajic\}@ufal.mff.cuni.cz}\\
\texttt{gauravjain@cse.iitb.ac.in}, \texttt{michal@hrusecky.net}, \texttt{michalisek@gmail.com}
}
%end author
%\title{Instructions for ACL-IJCNLP 2009 Proceedings}
%
%\author{First Author\\
%  Affiliation / Address line 1\\
%  Affiliation / Address line 2\\
%  {\tt email@domain}  \And
%  Second Author\\
%  Affiliation / Address line 1\\
%  Affiliation / Address line 2\\
%  {\tt  email@domain}}

\date{}

\begin{document}
\maketitle
\thispagestyle{fancy}

\begin{abstract}
We describe our attempt to improve on previous English to Hindi machine
translation results, using two open-source phrase-based MT systems: Moses and
Joshua. We use several approaches to morphological tagging: from automatic word
classes, through stem-suffix segmentation, to a POS tagger.
%We also experiment
%with factored language models.
We evaluate various combinations of training data
sets and other existing English-Hindi resources. To our knowledge, the BLEU
score we obtained is currently the best published result for the IIIT-TIDES
dataset.
\end{abstract}

\section{Introduction}
\label{sec:intro}

Machine translation is a challenging task and more so with significant
differences in word order of the languages in question and with the target
language explicitly marking more details in word forms than the source language
does. Precisely this holds for the English-Hindi pair we study.

We try to
explore the problems on several fronts: \Sref{sec:data} describes our
careful cleanup and a few additions to the training data. \Sref{morf} is devoted
to several additional variants of morphological representation of the target
side.
%and in \Sref{flm} we study the impact of using factored language models
%\parcite{bilmes:kirchhoff:factored:2003} instead of independent standard
%$n$-gram models for ensuring
%output coherence.
\Sref{joshua} evaluates the impact of using a hiearchical
instead of phrase-based model. \Sref{human} concludes our experiments by
providing a preliminary human evaluation of translation quality.

\section{Data}
\label{sec:data}

\def\microsection#1{{\bf #1.}}

\microsection{Tides}
\label{sec:tides}
A dataset originally collected for the DARPA-TIDES surprise-language contest in
2002, later refined at IIIT Hyderabad and provided for the NLP Tools Contest at
ICON 2008 \citep{nlptoolscontest:2008:summary}: 50K sentence pairs for training,
1K development and 1K test data (1~reference translation per sentence).

The corpus is a general domain dataset with news articles forming the greatest
proportion. It is aligned on sentence level, and tokenized to some extent.
We found the tokenization insufficient
%(e.g. \textit{anglo-american} would be 1
%token instead of 3, making the data sparseness problem more severe)
and ran our own tokenizer on top of it.
Cleaning was also necessary due to significant noise in the data, often caused
by misconversion of Latin script.

We used Tides as our primary dataset and all reported scores are measured on its
test data. However, note that due to the processing we applied to both training
and test data, our results are not directly comparable to the results of the
2008 NLP Tools Contest.\footnote{For instance, the Joshua BLEU score of a model
trained on Tides only, as we present it later in \Tref{tab:joshua}, is 12.27 on
the cleaned data, and 11.10 on the raw data; see also \Tref{tab:previous}.} We are happy
to share our cleaning tools to make the experiments reproducible.\footnote{\url{https://wiki.ufal.ms.mff.cuni.cz/pub-company:icon2009}; accept the certificate.}

\microsection{Daniel Pipes}
\label{sec:dpipes}
A journalist Daniel Pipes' website:\footnote{\url{http://www.danielpipes.org/}} limited-domain articles about the Middle East. 
The articles are originally written in English, many of them are translated to
up to 25 other languages, including Hindi (322 articles, 6,761 sentence
pairs).

\microsection{Emille}
\label{sec:emille}
EMILLE corpus \citep{baker:2002} consists of three components: monolingual,
parallel and annotated corpora.
%There are fourteen monolingual corpora,
%including both written and (for some languages) spoken data for fourteen South
%Asian languages.''
%: Assamese, Bengali, Gujarati, Hindi, Kannada, Kashmiri,
%Malayalam, Marathi, Oriya, Punjabi, Sinhala, Tamil, Telegu and Urdu.
%The EMILLE
%monolingual corpora contain approximately 92,799,000  words (including 2,627,000
%words of transcribed spoken data for Bengali, Gujarati, Hindi, Punjabi and
%Urdu).''
The parallel corpus consists of 200,000 words of text in English and
its accompanying translations in Hindi and other languages. Whenever we mention
Emille, we mean the parallel English-Hindi section.

The original Emille turned out to be very badly aligned and had spelling errors,
so we worked with a manually cleaned and aligned subset of 3,501 sentence
pairs.\footnote{We are very grateful to Om Dammani and his colleagues from IIT
Mumbai for making their corrected subset of Emille available to us.}

%\subsection{Other Data}
\label{sec:data-o}

We also tried various other small datasets:

 \microsection{ACL 2005}
 A subset of Emille, used in the shared task of the
 ACL 2005 workshop on ``Building and Using Parallel Texts: Data Driven Machine
 Translation and Beyond''.\footnote{Downloaded from the workshop website at
 \url{http://www.cse.unt.edu/~rada/wpt05/}. We used this dataset on the
 assumption that it might be better aligned, compared to the original Emille
 parallel corpus.}

 \microsection{Wiki NEs} 
 We have extracted English titles of Wikipedia
 entries, that are translations (or rather transcriptions) of Hindi, Marathi and
 Sanskrit named entities (names of persons, artifacts, places, etc.), and their
 translations in Devanagari.
% We used XML abstracts of English Wikipedia; the
% begining of an entry looks something like this, so it is easy to parse it with
% a regular expression:

%\hspace{1em} \textbf{Mumbai} (Marathi: \hindi{मुंबई}, \textit{Mumbaī}\ldots

% We took not only Hindi, but also Marathi and Sanskrit entities on the
% assumption that these are also commonly used in Hindi texts. See for example
% Mumbai (Bombay) above.

 \microsection{Shabdanjali}
 A GPL-licensed collaborative English-Hindi dictionary, containing about 26,000
 entries, available on the web.
% Available on the web in various formats and encodings. It is formatted
% similarly to printed bilingual dictionaries, so several filters had to be
% applied to refine a simple list of word pairs out of it.

 \microsection{Agriculture domain parallel corpus}
 English-Hindi-Marathi-UNL parallel corpus from Resource Center for Indian
 Language Technology
 Solutions.\footnote{\url{http://www.cfilt.iitb.ac.in/download/corpus/parallel/agriculture_domain_parallel_corpus.zip}}
 It contains 17,105 English and 13,248 Hindi words.

The impact of adding additional data to the Tides training data is partially
illustrated by \Tref{tab:joshua} and \Tref{tab:human:more}. Adding other data
has similar results.

\subsection{Normalization}

As always with statistical approaches, we want our training data to match the
test data as closely as possible. There are style variations throughout our data
that can and should be normalized automatically. For instance, the Tides corpus
usually (except for some conversion errors) terminates a sentence with the
period (``.''). However, some of our additional data sets use the traditional
Devanagari sign called \textit{danda} (``\hindi{।}'') instead. Our normalization
procedure replaces all dandas by periods, converts Devanagari digits to ASCII
digits and performs other minor reductions of the character set (e.g.
normalizing
non-ASCII punctuation). Some changes affect the English data as well.

\subsection{Lessons Learnt}

Hindi as the target language possesses some features that negatively influence
MT performance: richer morphology than English, greater noise in training data
and harder sparse-data problem due to vocabulary that combines words from
various etymological sources.

One common cause of data sparseness is unstable orthography of English and other
loanwords (or even transcribed citations), cf. the following counterparts of the
English word ``standards'', all present in the data:

\begin{compactitem}
\item \hindi{स्टैंडर्डज} \textit{(sṭaiṁḍarḍaja)} \translit{(sṭaiṁḍarḍaja)}
\item \hindi{स्टैंडर्डस} \textit{(sṭaiṁḍarḍasa)}
\item \hindi{स्टैंडर्ड्स} \textit{(sṭaiṁḍarḍsa)}
\end{compactitem}

Also common is the case where genuine English text has been (during some
processing at the site of the data provider) interpreted as Devanagari encoded
using Latin letters. Thus, \hindi{ईन्ङोर्मटिओन् छोम्मिसिओनेर्} \textit{(īnṅormaṭion
chommisioner)} should actually read (even in the Hindi text) \textit{Information
Commis(s)ioner}. If transcribed to Devanagari, it would be
% Zdá se, že font Madan neobsahuje nuktu ani znaky s ní.
% i+n+virám+f+o...
%\hindi{इन्फ़ोर्मेशन कोमिशनेर}
% i+n+virám+ph+nukta+o...
\hindi{इन्फ़ोर्मेशन कोमिशनेर}
\textit{(informeśana komiśanera)}.

\section{Target-Side Morphology}
\label{morf}
Richer morphology on the target side means that besides selecting the
\textit{lexically correct} Hindi equivalent of an English word (such as
\textit{book $\rightarrow$} \hindi{किताब} \textit{(kitāba)}), the system also
must correctly guess the \textit{grammatical features} (such as the direct vs.
oblique case in \textit{books $\rightarrow$} \hindi{किताबें} \textit{(kitābeṁ)} vs.
\hindi{किताबों} \textit{(kitāboṁ)}).
% Tu následující větu o shodě lze při nedostatku místa vynechat.
Moreover, grammatical agreement (such as
\hindi{मॆरी किताब} \textit{(merī kitāba)} ``my book'' vs. \hindi{मॆरा कमरा}
\textit{(merā kamarā)} ``my room'') further multiplies possible translations
of an English word.
A separate model of target-language morphology may help the system select the
correct translation.

We have tried two supervised and four unsupervised approaches to Hindi
morphology, see below.
%
%a full POS tagging
%with a morphology and POS tagger from IIT Mumbai (\Sref{sec:tagger}) and
%suffixes learned from a textbook (\Sref{sec:textbook}). We also tried several
%unsupervised methods. A short comparison of the morphological factors is given
%in the \Tref{tab:morph}.
%
The results in \Tref{tab:morph} and \Tref{tab:human} show that we were not able
to achieve any improvements by using real POS tags
%(see \Sref{sec:tagger})
compared to automatic word classes produced by \texttt{mkcls} or classes created
by hand.
%(\Sref{sec:textbook}).

\subsection{Supervised Morphology} 
\label{sec:pos}

\microsection{POS tagger}
\label{sec:tagger}
The only Hindi POS tagger we were able to find on the web, download and use is a part of the
GATE framework \citep{Cunningham:2002rw}. That is however more of a
demonstration, than a real tagger.
%Other than that, we found no usable tagger.
%That is probably one of the reasons that lead researchers to pursue
%unsupervised
%approaches to morphology; see e.g. \citet{BoStEnglishHindiTranslation2008}.
Fortunately, we also had the opportunity to work with a CRF-based tagger from IIT
Mumbai \citep{Kuhoo-Gupta:2006gd}.\footnote{Many thanks to Pushpak
Bhattacharyya for making this tool available to us.}
%dz 14.10.2009: následující URL možná už nefunguje?
%\footnote{\url{http://www.cfilt.iitb.ac.in/tools/Hindi_POStagger.zip}}

%The java application providing morphological analyses turned out to be a good
%detector of dirty input data, because it crashed on various characters that
%(mostly) did not belong to the input text. Most of these characters probably
%appeared in the input text because of failed transcription as described in
%\Sref{sec:tides}. However, since it crashes on occurrences of \% or \$, or
%square brackets, which morphology uses for its own purpose, it can pose a
%problem when these characters occur legitimately in the input text.

%Once we cleaned-up and normalized the input data, morphology and POS tagger ran smoothly.

\microsection{Textbook suffixes}
\label{sec:textbook}
Since the main morphological device of Hindi is alternating word suffixes,
one can model suffixes instead of morphological tags.
Our unsupervised approaches (see below) automatically acquire long but noisy
lists of Hindi suffixes, some of which may bear grammatical meaning.
As an alternative, we compiled a short list of 31 suffixes (including
duplicates) for regular declension of nouns, verb formation etc.
The cost of such a list was negligible: one of the authors (who does not speak
Hindi) spent less then two hours with a Hindi textbook
\citep{teach_yourself_hindi}, looking for grammar-describing sections.

\subsection{Unsupervised Morphology} 
We have tried two simple unsupervised methods:
\begin{itemize}
\item Automatic classes created by \texttt{mkcls}\footnote{\url{http://www.fjoch.com/mkcls.html}} tool contained in GIZA++ installation. 
\item simple n-character long suffixes
\end{itemize}
and two more elaborate methods:
%Affisix and Hindomor.

\label{sec:aff}
{\bf Affisix} is an open source tool for unsupervised recognition of affixes in a
given language. It takes a large list of words and generates a list
of possible suffixes or prefixes scored according to a selected method or a
combination of methods. In this case, we used it to obtain the 100 top scoring
suffixes
using backward entropy and backward difference entropy methods (see
\cite{affisixtsd2008} for details). The longest possible suffix seen in a word
was then treated as the morphological tag of the word, leaving some words with a
blank tag.

\label{sec:hindomor}
{\bf Hindomor} is another tool for unsupervised segmentation of words into
morphemes. It was originally published in the context of information
retrieval \citep{Zeman:2008}.

The tool has been trained on the word types of the Hindi side of the Tides
corpus. For every word the algorithm searches for positions where it can be cut
in two parts: the stem and the suffix. Then it tries to filter the stem and
suffix candidates so that real stems and suffixes remain. The core idea is that
real stems occur with multiple suffixes and real suffixes occur with multiple
stems.

Given the lists of stems and suffixes obtained during training, we want to find
the stem-suffix boundary in a word of the same language. Theoretically, we could
use the learned stem-suffix combinations to require that both stem and suffix be
known. However, this approach proved too restrictive, so we ended up in using
just the list of suffixes. If a word ends in a string equal to a known suffix,
the morpheme boundary is placed at the beginning of that substring.

\begin{table}[t]
\begin{center}
\small
\begin{tabular}{l  l | ll}
factor & BLEU &
factor & BLEU\\
\hline
tag & 12.03±0.75 &	hitbsuf & 11.58±0.74\\
wc50 & 11.97±0.73 &	hindomor2 & 11.55±0.74\\
wc10 & 11.76±0.74 &	hindomor1 & 11.54±0.71\\
lcsuf3 & 11.66±0.75 &	affddf & 11.50±0.7\\
lcsuf1 & 11.63±0.72 &	affbdf & 11.33±0.72\\
hindomor3 & 11.60±0.73 &	lcsuf2 & 11.14±0.74\\
\end{tabular}
\end{center}
\caption{Target side morphology: Using different additional factors for second
language model of MT system and its  effect on BLEU score. Trained on
IIIT-TIDES only. 
tag -- POS tags; wc\textit{n} -- \textit{n} word classes from mkcls; hitbsuf --
word classes created by hand; lcsuf\textit{n} -- simple
n-character suffixes; hindomor\textit{n};
aff\textit{xxx} -- Affisix
} 
\label{tab:morph}
\end{table}

The best BLEU score we achieved with Moses while making experiments with different
factors for target side morphology was 12.22±0.78:
trained on Tides and Daniel Pipes data with \texttt{hitbsuf}
as the additional factor.

\section{Hierarchical Phrase-Based Models}
\label{joshua}

One of the major differences between English and Hindi is the word order.
Massive reordering must take place during translation.
That is why we conducted several experiments with hierarchical
translation models \citep{chiang:2007}, namely with Joshua \citep{joshua:2009},
and compared the results with classical phrase-based models \citep[Moses, ][]{moses:2007}.

Hierarchical phrase-based translation is based on synchronous context-free grammars (SCFG).
Like phrase-based translation, pairs of corresponding source- and target-language phrases
(sequences of tokens) are learnt from training data. The difference is that in hierarchical
models, phrases may contain ``gaps'', represented by nonterminal symbols of the SCFG.
If a source phrase $f$ contains a nonterminal $X_1$, then its translation $e$ also
contains that nonterminal, and the decoder can replace the nonterminal by any phrase $f_1$
and its translation $e_1$, respectively. An illustrative rule for English-to-Hindi translation is

%Zatím neumíme hindštinu uvnitř matematiky, ale neškodilo by, kdybysme ji tam nějak dostali.
%$$X \rightarrow \langle X_1 of X_2 , X_2 \hindi{का} X_1 \rangle$$

\begin{center}
$X \rightarrow \langle X_1\cdot of\cdot X_2\rangle , \langle X_2\cdot \mathsf{का} \cdot X_1 \rangle$
\end{center}

where the Hindi word \hindi{का} \textit{(kā)} is one of several grammatical forms of the Hindi equivalent
of ``of'', and the subscripts on the nonterminals cause the two (noun) phrases around \textit{of}
to be reordered around \hindi{का} in the translation.

Hierarchical models have been known to produce better results than classical phrase-based
models \citep{chiang:etal:2005}.

In our experiments we used the Joshua implementation of the hierarchical phrase-based
algorithms.\footnote{Many thanks to the team at Johns Hopkins University for creating
Joshua and making it publicly available.} We set the maximum phrase length to 5,
MERT worked with 300 best translation hypotheses per iteration, and the number of
iterations was limited to 5.

Unless stated otherwise, our experiments use a trigram language model trained on the
target side of the Tides training data. In accord with \citet{BoStEnglishHindiTranslation2008},
we found additional out-of-domain language models damaging to BLEU score.

Symmetrical word alignments for grammar extraction were obtained using Giza++
and the scripts accompanying Moses. Alignments were computed on first four
(lowercased) characters of each training token, and the \textit{grow-diag-final-and}
symmetrization heuristic was used.

The results of the hierarchical models trained on various datasets,
compared with classical phrase-based models are shown in \Tref{tab:joshua}.

% - Popsat důležité varianty experimentů, které jsme zkoušeli:
% -- Oficiální testování Ondrovým počítadlem (Ondra se zaručil, že na tokenizaci to nesahá):
%    cp ~bojar/diplomka/granty/hindi/statmt/playground/exp.mert.SRCtides.train+en+lc.TGT+hi+for.c9ce7fbc.-fe.r0-0DEVtides.dev.Sbeam.20090806-1700/testbleu .
%    gunzip -c /net/projects/hindi/data/augmented_corpora/tides.test/hi.gz > test.hi
%    testbleu test.hi < /net/projects/hindi/experiments/joshua/josh11-tmtides-lmtides-lm3-preicon2-prenorm-pretokdz/eval/test.joshua.hi
% -- baseline: trénování na tides (srovnat s Mosesem letos i loni)
%    josh11-tmtides-lmtides-lm3-prebase = 11.28 (Ondra naměřil 11.10±0.79) (pozor, tenhle výsledek nemám s Ondrovým alignmentem lcstem4!)
%    josh11-tmtides-lmtides-lm3-preicon2-prenorm-pretokdz = 12.27 (po normalizaci - kruci, měli bychom ukázat i nějaký výsledek před normalizací, jinak naše tvrzení, že jsme lepší než kdokoli loni, není zajímavé)
%    testbleu: 12.27±0.83
%    Moses: 1*LM 2*tides.train                                                     	11.46±0.72
% -- trénování na tides a daniel pipes
%    ac-tmtides+dpipes-lmtides-lm3 (allcstem4, ale tehdy jsem to ještě do názvu pokusu nepsal)
%    = 12.58
%    testbleu: 12.58±0.77
%    Moses: 1*LM 2*tides.train 1*dani                                              	11.93±0.75
% -- trénování na tides, daniel pipes a emille
%    ac-allcstem4-tmtides+dpipes+omille-lmtides-lm3 = 11.32
%    testbleu: 11.32±0.74
%    Moses: 1*LM 2*tides.train 1*dani 1*emille-om                                  	10.06±0.72
% -- trénování na tides, daniel pipes a slovník (zkusit ještě ten opravený?!!!)
%    ac-allcstem4-tmtides+dpipes+dictfilt-lmtides-lm3 = 12.43 (to je starý neopravený a malý slovník)
%    testbleu: 12.43±0.79
%    Moses: 1*LM 2*tides.train 1*dani 1*dictfilt                                   	11.90±0.78
% -- ke všem číselným výsledkům v této kapitole by to chtělo hned vedle vidět odpovídající výsledky Mosese.
% PS: V sekci o Joshuovi by se mi libilo, kdybys to nejak propojil s lonskym reorderingek: ze vlastne podstata SCFG je svym zpusobem syntakticky transfer, tedy to, co jsme vloni hrube resili reorderingovymi pravidly.

\begin{table}[ht]
\begin{centering}
\begin{tabular}{l|l|l}
\textbf{Parallel data} & \textbf{Joshua} & \textbf{Moses} \\
\hline
Tides & 12.27±0.83 & 11.46±0.72\\
Tides+DP & \textbf{12.58±0.77} & 11.93±0.75\\
Tides+DP+Emille & 11.32±0.74 & 10.06±0.72\\
Tides+DP+Dict & 12.43±0.79 & 11.90±0.78\\
\end{tabular}
\caption{Results of Joshua compared with Moses}
\label{tab:joshua}
\end{centering}
\end{table}

\section{Results}
\label{results}

As far as automatic evaluation is concerned, the best result reported on in
this paper is the 12.58 BLEU of Joshua trained on Tides and Daniel Pipes
(\Tref{tab:joshua}). Moses was not able to outperform this score despite
its ability to learn factored models. The best Moses score is 12.22 (morphology).
%(in
%tandem with the Mumbai tagger, see \Tref{tab:morph})

The greatest mystery is the fact that adding Emille to the training data
does not improve results with either system.
% Joshua: tides+dpipes+omille: 0.1298 dev, 0.1132 test.
% Joshua/Emille/Gaurav:        0.1285 dev, 0.1045 test.
% Bez omilla (tides+dpipes):   0.1124 dev, 0.1258 test.
Although we are not able to explain this in full, we made a few observations:
%\begin{itemize}
%  \item

{\bf 1.}
When asked to extract its SFCG, Joshua needs to see the source
        (English) side of the data to be decoded, and extracts only the rules
        relevant to the dataset. Thus we extract one grammar for the
        development data (used for minimum error rate training) and another
        for the test data. While the development grammar changes when Emille
        is added, the test grammar remains the same. As if Emille was unable
        to approximate the test data in any way.
%  \item

{\bf 2.}
MERT optimizes 5 feature weights of Joshua: the language model
        probability, the word penalty (preference for shorter translations),
        and three translation features: $P(e|f)$, $P_{lex}(f|e)$ and
        $P_{lex}(e|f)$. When Emille is involved, MERT always pushes the
        non-lexical translation probability extraordinarily high, and causes
        overfitting. While for other experiments we usually saw better BLEU
        scores on test data than on development data, the opposite was the case
        with Emille.
%\end{itemize}

\section{Related Research}

% Tohle zamlcime, stejne jsme v tabulce.
%\perscite{BoStEnglishHindiTranslation2008} tried to improve a statistical
%phrase-based baseline MT
%system by adding more data, standard lexicalized and a heuristical rule-based
%reordering and an explicit modeling of (unsupervised) morphological coherence
%on the target side. Neither of the approaches helped much.

\perscite{pushpak:2009} improve Hindi morphology by transfering additional
information from English parse tree and semantic roles, in addition to some
hard-coded syntax-based reordering. However, they use an unspecified corpus
making their results incomparable.

% XXX reorderingy? no jen jednoho zastupce, a jako ze Joshua leccos resi sam

%\subsection{Previous Work}
\label{previous}

The ICON 2008 NLP Tools Contest \citep{nlptoolscontest:2008:summary}
included translation of Tides test data. Although our retokenized and cleaned
data make a direct comparison impossible, our preliminary experiments on the
unclean data are comparable. In \Tref{tab:previous} we compare four results
presented at ICON 2008 with our hierarchical model trained on (unclean) Tides,
with a Tides trigram language model.

\begin{table}[ht]
\begin{centering}
\begin{tabular}{l|r}
\textbf{System} & \textbf{BLEU} \\
\hline
Mumbai \citep{Damani:2008} & 8.53\\
Kharagpur \citep{goswami:2008} & 9.76\\
Prague \citep{BoStEnglishHindiTranslation2008} & 10.17\\
Dublin \citep{srivastava:2008} & 10.49\\
present Joshua & 11.10\\
\end{tabular}
\caption{Previous work compared to our hierarchical model (Joshua) on unclean data}
\label{tab:previous}
\end{centering}
\end{table}

\section{Human Evaluation}
\label{human}

The richer morphology and a high degree of reordering are known to render BLEU
unreliable. While we still optimize for BLEU, our manual analysis reveals
relatively low correlation. An extensive study similar to
\cite{callisonburch-EtAl:2009:WMT} would be very valuable.

We were able to conduct three small-scale manual evaluations. Our
annotator was given the source English text and four or five Hindi translations in a
randomized order of 100
sentences (each time a different subset of the full test set). He assigned a
mark to each of the four Hindi translations: giving no
mark  (``0'') indicates a completely incomprehensible translation. A single star
(``*'') denotes sentences with severe errors and hard to understand but still
related to the input. Two stars (``**'') are assigned to more or less
acceptable translations, possibly with many errors but understandable and
conveying most of the original meaning.

\Tref{tab:human} evaluates some basic configurations of Moses. For a comparison,
we include the reference translations (REF) among the hypotheses (due to the
randomization, the annotator cannot be entirely sure which of the hypotheses is
the reference but often it can be guessed from the striking difference in
quality). We see that even 6 reference translations were marked as inadequate
and 11 as very bad. Because the test set is a part of the Tides corpus, these
figures give a rough estimate of the overall corpus quality.

The remaining figures in \Tref{tab:human} document that (while somewhat
suspicious by itself), the Tides training data are most informative with respect
to the (Tides) test set: a system trained completely out-of-domain on everything
except Tides was able to deliver only 3 acceptable translations (OOD). The TIDP
and WC10 systems compare the effect of adding more parallel data (TIDP for
Tides+DanielPipes) vs. adding the unsupervised morphological factor (WC10, 10
word classes). We observe that the difference indicated by human evaluation is
rather convincing: nearly twice as many acceptable sentences in TIDP, but
negligible in BLEU. This confirms our doubts about the utility of BLEU scores
for languages with rich morphology and the neccessity to regularly run manual
evaluations.

\begin{table}[ht]
\begin{centering}
\begin{tabular}{l|r|r|r|r}
{\bf System} & {\bf 0} & {\bf*} & {\bf**} & {\bf BLEU}\\
\hline
REF  &  6    &  11  &   83   &  -- \\
OOD &   80  &   17   &  3   &   1.85±0.24\\
TIDP &  26  &   44   &  30   &  11.93±0.75\\
WC10 &  38   &  46   &  16   &  11.76±0.74\\
\end{tabular}
\caption{Manual evaluation of some basic Moses setups: training out of domain
(OOD) and adding either more parallel data (TIDP) or a factor for morphological
coherence (WC10) compared to the quality of the reference translation (REF).
}
\label{tab:human}
\end{centering}
\end{table}

In \Tref{tab:human:mojo}, we evaluate the difference between phrase-based
(Moses) and hierarchical (Joshua) translation model. Both systems are trained in
identical conditions: both the translation and the language model are trained on
Tides and DanielPipes. For Moses, we wanted to confirm the observation that more
data are more important than ensuring morphological coherence on another 100
manually judged sentences. Therefore, we report also Moses-DPipes+POStags, a
setup trained on Tides only but including the supervised morphology of Mumbai
tagger.

We see that while the BLEU scores indicate the superiority of the hierarchical
model over the phrase-based model with lexicalized reordering, the difference in
manual judgments seems less convincing. Only 3
sentences were ranked better. On the other hand, we confirm that even the
supervised morphology is less important than a good parallel data source.

\begin{table}[ht]
\begin{small}
\begin{centering}
\begin{tabular}{l|r|r|r|r}
{\bf System} & {\bf 0} & {\bf*} & {\bf**} & {\bf BLEU}\\
\hline
REF   &  6   &   10    &  84   &  -- \\ 
Joshua  &  32   &  37   &  31   &   12.58±0.77\\
Moses  &  35  &   35   &  30   &  11.93±0.75\\
%Moses+DPipesR  &  33  &   42  &   25  &   11.80±0.73\\
% Moses+DPipesR je reduced English, zatajuju
Moses-DPipes+POStags  & 32  &   42   &  26   &  12.03±0.75\\
\end{tabular}
\caption{%
Manual judgements of hierarchical (Joshua) vs. phrase-based (Moses) translation.
}
\label{tab:human:mojo}
\end{centering}
\end{small}
\end{table}

\Tref{tab:human:more} provides manual analysis of our baseline Moses setup
(simple
phrase-based translation, lexicalized reordering, language model trained on the
full target side of the parallel data) trained on various subsets of
our parallel data. We start with the combination of Tides and DanielPipes
(TIDP), add Emille (EM), all other corpus sources (oth) and finally the
dictionary Shabdanjali in two variants: full (DICTFull) and filtered to contain
only Hindi words confirmed in a big monolingual corpus (DICTFilt). The BLEU
scores indicate that Emille hurts the performance when tested on Tides test
set. This surprising result was indeed confirmed by the manual analysis: only 12
instead of 19 sentences were translated acceptably. Adding further data reduces
the detrimental effect of Emille (not observed in BLEU scores) but the best
performance is achieved by
Tides+DanielPipes only.  Note that this final manual evaluation was based on a
smaller dataset of 53 sentences only.

\begin{table}[ht]
\begin{small}
\begin{centering}
\begin{tabular}{l|r|r|r|r}
{\bf System} & {\bf 0} & {\bf *} & {\bf**} & {\bf BLEU}\\
\hline
REF         &  0 &          8   &   45   & -- \\
TIDP           &  20 &       14   &  19   &11.89±0.76\\
TIDPEM        &   22 &          19   &   12   &9.61±0.75\\
TIDPEMoth      &   17 &         25  &    11   & 10.97±0.79\\
TIDPEMothDICTFilt     &    23 &  17  &    13  & 10.96±0.75\\
TIDPEMothDICTFull    &    22 &   16  &    15  & 10.89±0.69 \\
\end{tabular}
\caption{The effect of additional (out-of-domain) parallel data in phrase-based
translation.}
\label{tab:human:more}
\end{centering}
\end{small}
\end{table}

\section{Conclusions and Future Work}
\label{sec:concl}
%Do závěru bychom mohli plácnout něco jako ``We believe that most techniques
%described in this paper are straightforwardly generalizable to other Indian SOV
%languages.''

We have tried several ways to improve state-of-the-art in English-Hindi SMT,
however results are mixed:

There are quite a few strange results: 
\begin{itemize}
\item POS tagging does not give better results than automatic word classes.
Hindi textbook-based manual word classes were even better.
\item Adding more training data to Tides either helps insignificantly, or (more
often) hurts BLEU score. 
\item BLEU may not correlate with human judgement. Our limited experiments show
that adding training data may hurt BLEU but improve quality by human judgement.
%\item It has been shown before by \citet{BoStEnglishHindiTranslation2008} that 
%big monolingual Hindi data for LM didn't help either.
\end{itemize}

In our future work we want to further explore problems with existing datasets,
the use of morphology, and the relation of output quality measured in terms of
BLEU vs.
human
judgement. We also believe that there is room for improvement in the quality and
amount of available parallel data.

On the other hand, we have shown that our hand-crafted word classes and some
additional data help Moses achieve significantly better results than reported
previously. Hierarchical decoder Joshua can capture word order even better than
Moses. Its results are always slightly better. And as far as we know, our
current results are the best that have been reported on this dataset.

%\XXX -- někde zmínit, že "čištění" (ve smyslu Mosese, tedy dlouhých vět)
%výsledky kazí, protože nám zmenšuje trénovací data. O to víc je vlastně
%zvláštní, že přidání trénovacích dat typu Emille výsledky taky kazí.

\begin{small}
\bibliography{paper}
\end{small}

\end{document}

### 2009-10-27 aclass.pl

@@ -19,6 +19,7 @@ binmode(STDOUT, ":utf8");
 binmode(STDERR, ":utf8");
 use Getopt::Long;
 use csts;
+use lib '/home/zeman/projekty/parser';
 use strom;
 
 
@@ -48,6 +49,8 @@ $n_spravne = 0 if($n_spravne eq "");
 $n_spatne = $n_celkem-$n_spravne;
 $uspesnost = $n_celkem ? $n_spravne/$n_celkem : 0;
 print STDERR ("A $n_celkem - G $n_spravne - B $n_spatne - P $uspesnost\n");
+print STDERR ("Neznámá dvojice mznaček, ale známá mznačka dítěte: $n_neznama_dvojice        z toho chyb $n_chyb_neznama_dvojice\n");
+print STDERR ("Není známa ani mznačka dítěte:                     $n_neznama_dvojice_i_dite z toho chyb $n_chyb_neznama_dvojice_i_dite\n");
 
 
 
@@ -69,24 +72,41 @@ sub zpracovat_vetu
         # Odstranit z klíče případné tabulátory a konce řádků, protože totéž jsme dělali při tréninku.
         $klic =~ s/[\t\r\n]+/ /sg;
         my $hodnota = $stat{$klic};
+        my $neznama_dvojice = 0;
+        my $neznama_dvojice_i_dite = 0;
         # Jestliže jsme nenašli žádnou sznačku pro danou dvojici mznaček, zkusíme najít sznačku pro mznačku dítěte.
         if($hodnota eq "")
         {
             $klic = $slovo->{znacka};
             $klic =~ s/[\t\r\n]+/ /sg;
             $hodnota = $stat{$klic};
+            $n_neznama_dvojice++;
+            $neznama_dvojice = 1;
         }
         # Jestliže jsme ani teď nenašli žádnou sznačku, zkusíme najít nejčastější sznačku.
         if($hodnota eq "")
         {
             $klic = "";
             $hodnota = $stat{$klic};
+            $n_neznama_dvojice_i_dite++;
+            $neznama_dvojice_i_dite = 1;
         }
         # Zkontrolovat, zda hodnota z naší statistiky odpovídá skutečné syntaktické značce slova.
         $slovo->{afun} =~ s/[\t\r\n]+/ /sg;
         if($hodnota eq $slovo->{afun})
         {
             $n_spravne++;
+        }
+        else
[…]

### 2009-10-27 parse.pl

@@ -38,6 +38,8 @@ $starttime = time();
 my $inisoubor = "parser.ini"; # jméno souboru s konfigurací
 # parse.pl --i parser2.ini
 GetOptions('model=s' => \$model, 'ini=s' => \$inisoubor);
+# Výchozí nastavení parametrů.
+%konfig = parse::vychozi_konfig();
 parse::precist_konfig($inisoubor, \%konfig);
 if($model ne "")
 {
@@ -270,6 +272,10 @@ sub cist_statistiku
         my $k = $1;
         my $c = $2;
         my $udalost = $k;
+        # Pokud původní popis události obsahoval tabulátor, byl tabulátor zakódován a musíme ho teď dekódovat.
+        # Kvůli kódování tabulátorů byly zakódovány i ampersandy.
+        $udalost =~ s/&tab;/\t/g;
+        $udalost =~ s/&amp;/&/g;
         $statref->{$udalost} = $c;
         # Přičíst i do celkového počtu všech událostí (jmenovatel).
         $celkem += $c;

### 2009-10-27 train.pl

@@ -20,6 +20,8 @@ $starttime = time();
 my $inisoubor = "parser.ini"; # jméno souboru s konfigurací
 # train.pl --i parser2.ini
 GetOptions('ini=s' => \$inisoubor);
+# Výchozí nastavení parametrů.
+%konfig = parse::vychozi_konfig();
 parse::precist_konfig($inisoubor, \%konfig);
 # Nastavit, který výstup půjde na STDOUT. Ostatní půjdou na STDERR.
 $vystupy::vystupy{stat}{stdout} = 1;
@@ -46,7 +48,8 @@ $konfig{hook_zacatek_cteni} = sub
     vypsat("prubeh", "Maska pro jména souborů s daty: $maska\n");
     vypsat("prubeh", "Nalezeno ".($#{$soubory}+1)." souborů.\n");
 };
-csts::projit_data($konfig{train}, \%konfig, \&zpracovat_vetu);
+$tmpfile = zkopirovat_vstup_do_tmp($konfig{train});
+csts::projit_data($tmpfile, \%konfig, \&zpracovat_vetu);
 vypsat("prubeh", "Počet vět  = $veta\n");
 vypsat("prubeh", "Počet slov = $slovo\n");
 # Teď ještě natrénovat modely n-tic. Nemohli jsme to dělat všechno při jednom
@@ -58,9 +61,10 @@ if($konfig{ntice})
     $veta = 0;
     $slovo = 0;
     $ohlasena_veta = 0;
-    csts::projit_data($konfig{train}, \%konfig, \&zpracovat_vetu_ntice);
+    csts::projit_data($tmpfile, \%konfig, \&zpracovat_vetu_ntice);
     ntice::vypsat_do_stat();
 }
+unlink($tmpfile);
 # Poslat mi mail, že trénink je u konce. Musíme do mailu dát nějaký existující
 # soubor. Stačil by mi sice prázdný mail jen s předmětem zprávy, ale pokud bych
 # k tomu chtěl využít existující mechanismy, vznikl by mi tím na disku prázdný
@@ -727,7 +731,11 @@ sub ulozit
     $n = 1 if($n==0); # kvůli dělení při hlášení pokroku
     for(my $i = 0; $i<=$#stat; $i++)
     {
-        vypsat("stat", "$stat[$i]\t$statref->{$stat[$i]}\n");
+        # Popis události nesmí obsahovat tabulátor, jinak by statistika nešla opět načíst.
+        my $ud = $stat[$i];
[…]

### 2009-10-27 papers/2009-icon-hyderabad/paper.tex
Clanek aktualizovan, dosud to byla spis kopie clanku o prekladu.

@@
 \setmathsf{Lohit Hindi}%\XXX
-\newfontinstance\hi[Script=Devanagari]{Lohit Hindi}
-\newcommand{\hindi}[1]{{\hi #1}}
-\newfontinstance\tl{Gentium}
-\newcommand{\translit}[1]{{\tl \textit{#1}}}
+%\newfontinstance\hi[Script=Devanagari]{Lohit Hindi}
+\newfontinstance\hifont[Script=Devanagari]{Code2000}
+\newfontinstance\bnfont[Script=Bengali]{Code2000}
+\newfontinstance\tefont[Script=Telugu]{Code2000}
+\newfontinstance\translitfont{Gentium}
+\newcommand{\hi}[1]{{\hifont #1}}
+\newcommand{\bn}[1]{{\bnfont #1}}
+\newcommand{\te}[1]{{\tefont #1}}
+\newcommand{\translit}[1]{{\translitfont \textit{#1}}}
 
@@
 %ps: Did you mean \citep and \citet (in-parentheses and textual reference)? There is even more, see `texdoc natbib`.
-%obo: your def seems right.
 \def\Sref#1{Section~\ref{#1}}
@@
 
-\title{English-Hindi Translation -- Obtaining Mediocre Results\\with Bad Data and
-Fancy Models%
-\thanks{ \hspace{.6em}The research has been supported by the grants GAAV ČR 1ET201120505 (Grant Agency of the Academy of Sciences of the Czech Republic), 
-MSM0021620838 (Czech Ministry of Education),
-FP7-ICT-2007-3-231720 (EuroMatrix Plus), and GAUK 4307/2009 (Grant agency of the Charles University).
-}
+\def\microsection#1{{\bf #1.}}
+
+\title{Maximum Spanning Malt: Hiring World's Leading Dependency Parsers to Plant Indian Trees%
+%\thanks{ \hspace{.6em}The research has been supported by the grant 
+%MSM0021620838 (Czech Ministry of Education).}
 }
 
-% oboXXX nedat Gaurava az za Michaly?
-% dz: jsem spíš proti. Konečný přínos Michalů je možná větší, ale pro Gaurava to byla ty 3 měsíce hlavní náplň činnosti a dělal přímo na samotném překladu, ne na "doplňcích".
-\author{Ondřej Bojar, Pavel Straňák, Daniel Zeman, Gaurav Jain, Michal Hrušecký, Michal Richter, Jan Hajič\\
-Univerzita Karlova v Praze, Ústav formální a aplikované lingvistiky,\\
-Malostranské náměstí 25, 118 00, Praha, Czech Republic\\ 
-\texttt{\{bojar,stranak,zeman,hajic\}@ufal.mff.cuni.cz}\\
-\texttt{gauravjain@cse.iitb.ac.in}, \texttt{michal@hrusecky.net}, \texttt{michalisek@gmail.com}
+% Tady je posuzování taky slepé? Až odtajním autora, nezapomenout odtajnit i grant!
+\author{%Daniel Zeman\\
+%Univerzita Karlova v Praze, Ústav formální a aplikované lingvistiky\\
+%Malostranské náměstí 25, CZ-11800, Praha, Czechia\\ 
+%\texttt{zeman@ufal.mff.cuni.cz}
 }
-%end author
+
 %\title{Instructions for ACL-IJCNLP 2009 Proceedings}
@@
 \begin{abstract}
-We describe our attempt to improve on previous English to Hindi machine
-translation results, using two open-source phrase-based MT systems: Moses and
-Joshua. We use several approaches to morphological tagging: from automatic word
-classes, through stem-suffix segmentation, to a POS tagger.
-%We also experiment
-%with factored language models.
-We evaluate various combinations of training data
-sets and other existing English-Hindi resources. To our knowledge, the BLEU
-score we obtained is currently the best published result for the IIIT-TIDES
-dataset.
+We present our system used for participation in the ICON 2009 NLP Tools Contest: dependency parsing of Hindi, Bangla and Telugu.
+The system consists of three existing, freely available dependency parsers, two of which (MST and Malt) have been known to produce state-of-the-art structures on data sets for other languages.
+Various settings of the parsers are explored in order to adjust them for the three Indian languages, and a voting approach is used to combine them into a superparser.
+Since there is nothing novel about the approach used, substantial part of the paper is devoted to the analysis of errors the system makes on the given data sets.
 \end{abstract}
@@
 
-Machine translation is a challenging task and more so with significant
-differences in word order of the languages in question and with the target
-language explicitly marking more details in word forms than the source language
-does. Precisely this holds for the English-Hindi pair we study.
+Introduction goes here.
+\XXX Sections: Brief system description, Error analysis, Results, Conclusion
 
-We try to
-explore the problems on several fronts: \Sref{sec:data} describes our
-careful cleanup and a few additions to the training data. \Sref{morf} is devoted
-to several additional variants of morphological representation of the target
-side.
-%and in \Sref{flm} we study the impact of using factored language models
-%\parcite{bilmes:kirchhoff:factored:2003} instead of independent standard
-%$n$-gram models for ensuring
-%output coherence.
-\Sref{joshua} evaluates the impact of using a hiearchical
-instead of phrase-based model. \Sref{human} concludes our experiments by
-providing a preliminary human evaluation of translation quality.
+\section{System Description}
+\label{sec:system}
 
-\section{Data}
-\label{sec:data}
+Several good trainable dependency parsers have emerged during the past five years.
+The CoNLL-X \citep{buchholz-marsi:2006:CoNLL-X} and CoNLL 2007 \citep{nivre-EtAl:2007:EMNLP-CoNLL2007} shared tasks in multilingual dependency parsing have greatly contributed to the development of the parsers.
+Some of the parsers are now freely available on the web, some are even open-source.
+We selected three of the publicly available parsers for our experiments:
 
-\def\microsection#1{{\bf #1.}}
+%\microsection{MST Parser}
+\subsection{MST Parser}
+\label{sec:mst}
+The Maximum Spanning Tree (MST) parser \citep{mst} views the sentence as an oriented complete graph with edges weighted by a feature scoring function.
+It finds for the graph a spanning tree that maximizes the weights of the edges.
+A multi-class classification algorithm called MIRA is used to compute the scoring function.
 
+MST Parser achieved the best unlabeled attachment scores (UAS) for 9 out of the 13 languages of CoNLL-X, and second best scores in two others.
+Parsing is fast but training the parser takes many hours on large treebanks.
+On small data however, multiple quick experiments with different settings are still doable.
+The parser is implemented in Java and freely available for download.\footnote{\url{http://sourceforge.net/projects/mstparser/}}
+
+%\microsection{Malt Parser}
+\subsection{Malt Parser}
+\label{sec:malt}
+The Malt Parser  \citep{malt} is a deterministic shift-reduce parser where input words can be either put to the stack or taken from the stack and combined to form a dependency.
+The decision, which operation to perform, is made by an oracle based on various features of the words in the input buffer and the stack.
+The default machine learning algorithm used to train the oracle is a sort of SVN (support vector machine) classifier \citep{svm}.
+
+Malt Parser has participated in both CoNLL-X and CoNLL 2007 shared tasks, and although it achieved the best UAS in three languages only, it usually scored among the five best parsers, sometimes with statistically insignificant difference from the winner.
+Malt Parser is really fast and its new Java implementation is open-source, freely available for download.\footnote{\url{http://maltparser.org/}}
+
+%\microsection{DZ Parser}
+\subsection{DZ Parser}
+\label{sec:dz}
+In order to combine the two above parsers, we needed a third parser.
+We picked DZ Parser \citep{dzparser}, which is also reasonably fast and freely available.\footnote{\url{http://ufal.mff.cuni.cz/~zeman/projekty/parser/}} Although its accuracy, if compared to MST or Malt, is worse by a wide margin, this parser proved useful because its only role was to help to form a majority whenever MST and Malt disagreed.
+
+\XXX Acknowledgements: poděkovat Joakimovi a Ryanovi za zveřejnění parserů.
+\XXX other dependency parsers available for download include the isbn parser by titov
 \microsection{Tides}
 \label{sec:tides}
-A dataset originally collected for the DARPA-TIDES surprise-language contest in
-2002, later refined at IIIT Hyderabad and provided for the NLP Tools Contest at
-ICON 2008 \citep{nlptoolscontest:2008:summary}: 50K sentence pairs for training,
-1K development and 1K test data (1~reference translation per sentence).
-
-The corpus is a general domain dataset with news articles forming the greatest
-proportion. It is aligned on sentence level, and tokenized to some extent.
-We found the tokenization insufficient
-%(e.g. \textit{anglo-american} would be 1
-%token instead of 3, making the data sparseness problem more severe)
-and ran our own tokenizer on top of it.
-Cleaning was also necessary due to significant noise in the data, often caused
-by misconversion of Latin script.
-
-We used Tides as our primary dataset and all reported scores are measured on its
-test data. However, note that due to the processing we applied to both training
-and test data, our results are not directly comparable to the results of the
-2008 NLP Tools Contest.\footnote{For instance, the Joshua BLEU score of a model
-trained on Tides only, as we present it later in \Tref{tab:joshua}, is 12.27 on
-the cleaned data, and 11.10 on the raw data; see also \Tref{tab:previous}.} We are happy
+We are happy
 to share our cleaning tools to make the experiments reproducible.\footnote{\url{https://wiki.ufal.ms.mff.cuni.cz/pub-company:icon2009}; accept the certificate.}
@@
 
-\microsection{Emille}
-\label{sec:emille}
-EMILLE corpus \citep{baker:2002} consists of three components: monolingual,
-parallel and annotated corpora.
-%There are fourteen monolingual corpora,
-%including both written and (for some languages) spoken data for fourteen South
-%Asian languages.''
-%: Assamese, Bengali, Gujarati, Hindi, Kannada, Kashmiri,
-%Malayalam, Marathi, Oriya, Punjabi, Sinhala, Tamil, Telegu and Urdu.
-%The EMILLE
-%monolingual corpora contain approximately 92,799,000  words (including 2,627,000
-%words of transcribed spoken data for Bengali, Gujarati, Hindi, Punjabi and
-%Urdu).''
-The parallel corpus consists of 200,000 words of text in English and
-its accompanying translations in Hindi and other languages. Whenever we mention
-Emille, we mean the parallel English-Hindi section.
-
-The original Emille turned out to be very badly aligned and had spelling errors,
-so we worked with a manually cleaned and aligned subset of 3,501 sentence
-pairs.\footnote{We are very grateful to Om Dammani and his colleagues from IIT
-Mumbai for making their corrected subset of Emille available to us.}
-
-%\subsection{Other Data}
-\label{sec:data-o}
-
-We also tried various other small datasets:
-
- \microsection{ACL 2005}
- A subset of Emille, used in the shared task of the
- ACL 2005 workshop on ``Building and Using Parallel Texts: Data Driven Machine
- Translation and Beyond''.\footnote{Downloaded from the workshop website at
- \url{http://www.cse.unt.edu/~rada/wpt05/}. We used this dataset on the
- assumption that it might be better aligned, compared to the original Emille
- parallel corpus.}
-
- \microsection{Wiki NEs} 
- We have extracted English titles of Wikipedia
- entries, that are translations (or rather transcriptions) of Hindi, Marathi and
- Sanskrit named entities (names of persons, artifacts, places, etc.), and their
- translations in Devanagari.
-% We used XML abstracts of English Wikipedia; the
-% begining of an entry looks something like this, so it is easy to parse it with
-% a regular expression:
-
-%\hspace{1em} \textbf{Mumbai} (Marathi: \hindi{मुंबई}, \textit{Mumbaī}\ldots
-
-% We took not only Hindi, but also Marathi and Sanskrit entities on the
-% assumption that these are also commonly used in Hindi texts. See for example
-% Mumbai (Bombay) above.
-
- \microsection{Shabdanjali}
- A GPL-licensed collaborative English-Hindi dictionary, containing about 26,000
- entries, available on the web.
-% Available on the web in various formats and encodings. It is formatted
-% similarly to printed bilingual dictionaries, so several filters had to be
-% applied to refine a simple list of word pairs out of it.
-
- \microsection{Agriculture domain parallel corpus}
- English-Hindi-Marathi-UNL parallel corpus from Resource Center for Indian
- Language Technology
- Solutions.\footnote{\url{http://www.cfilt.iitb.ac.in/download/corpus/parallel/agriculture_domain_parallel_corpus.zip}}
- It contains 17,105 English and 13,248 Hindi words.
-
-The impact of adding additional data to the Tides training data is partially
-illustrated by \Tref{tab:joshua} and \Tref{tab:human:more}. Adding other data
-has similar results.
-
 \subsection{Normalization}
 
-As always with statistical approaches, we want our training data to match the
-test data as closely as possible. There are style variations throughout our data
-that can and should be normalized automatically. For instance, the Tides corpus
-usually (except for some conversion errors) terminates a sentence with the
-period (``.''). However, some of our additional data sets use the traditional
-Devanagari sign called \textit{danda} (``\hindi{।}'') instead. Our normalization
-procedure replaces all dandas by periods, converts Devanagari digits to ASCII
-digits and performs other minor reductions of the character set (e.g.
-normalizing
-non-ASCII punctuation). Some changes affect the English data as well.
-
-\subsection{Lessons Learnt}
-
-Hindi as the target language possesses some features that negatively influence
-MT performance: richer morphology than English, greater noise in training data
-and harder sparse-data problem due to vocabulary that combines words from
-various etymological sources.
-
-One common cause of data sparseness is unstable orthography of English and other
-loanwords (or even transcribed citations), cf. the following counterparts of the
-English word ``standards'', all present in the data:
+blabla
 
 \begin{compactitem}
-\item \hindi{स्टैंडर्डज} \textit{(sṭaiṁḍarḍaja)} \translit{(sṭaiṁḍarḍaja)}
-\item \hindi{स्टैंडर्डस} \textit{(sṭaiṁḍarḍasa)}
-\item \hindi{स्टैंडर्ड्स} \textit{(sṭaiṁḍarḍsa)}
+\item \hi{स्टैंडर्डज} \textit{(sṭaiṁḍarḍaja)} \translit{(sṭaiṁḍarḍaja)}
+\item \hi{स्टैंडर्डस} \textit{(sṭaiṁḍarḍasa)}
+\item \hi{स्टैंडर्ड्स} \textit{(sṭaiṁḍarḍsa)}
+\item \bn{সরিয়ে} Tohle je bengálsky.
+\item \te{వాడుతున్నాం} Tohle je telugsky.
 \end{compactitem}
-
-Also common is the case where genuine English text has been (during some
-processing at the site of the data provider) interpreted as Devanagari encoded
-using Latin letters. Thus, \hindi{ईन्ङोर्मटिओन् छोम्मिसिओनेर्} \textit{(īnṅormaṭion
-chommisioner)} should actually read (even in the Hindi text) \textit{Information
-Commis(s)ioner}. If transcribed to Devanagari, it would be
-% Zdá se, že font Madan neobsahuje nuktu ani znaky s ní.
-% i+n+virám+f+o...
-%\hindi{इन्फ़ोर्मेशन कोमिशनेर}
-% i+n+virám+ph+nukta+o...
-\hindi{इन्फ़ोर्मेशन कोमिशनेर}
-\textit{(informeśana komiśanera)}.
-
-\section{Target-Side Morphology}
-\label{morf}
-Richer morphology on the target side means that besides selecting the
-\textit{lexically correct} Hindi equivalent of an English word (such as
-\textit{book $\rightarrow$} \hindi{किताब} \textit{(kitāba)}), the system also
-must correctly guess the \textit{grammatical features} (such as the direct vs.
-oblique case in \textit{books $\rightarrow$} \hindi{किताबें} \textit{(kitābeṁ)} vs.
-\hindi{किताबों} \textit{(kitāboṁ)}).
-% Tu následující větu o shodě lze při nedostatku místa vynechat.
-Moreover, grammatical agreement (such as
-\hindi{मॆरी किताब} \textit{(merī kitāba)} ``my book'' vs. \hindi{मॆरा कमरा}
-\textit{(merā kamarā)} ``my room'') further multiplies possible translations
-of an English word.
-A separate model of target-language morphology may help the system select the
-correct translation.
-
-We have tried two supervised and four unsupervised approaches to Hindi
-morphology, see below.
-%
-%a full POS tagging
-%with a morphology and POS tagger from IIT Mumbai (\Sref{sec:tagger}) and
-%suffixes learned from a textbook (\Sref{sec:textbook}). We also tried several
-%unsupervised methods. A short comparison of the morphological factors is given
-%in the \Tref{tab:morph}.
-%
-The results in \Tref{tab:morph} and \Tref{tab:human} show that we were not able
-to achieve any improvements by using real POS tags
-%(see \Sref{sec:tagger})
-compared to automatic word classes produced by \texttt{mkcls} or classes created
-by hand.
-%(\Sref{sec:textbook}).
-
-\subsection{Supervised Morphology} 
-\label{sec:pos}
-
-\microsection{POS tagger}
-\label{sec:tagger}
-The only Hindi POS tagger we were able to find on the web, download and use is a part of the
-GATE framework \citep{Cunningham:2002rw}. That is however more of a
-demonstration, than a real tagger.
-%Other than that, we found no usable tagger.
-%That is probably one of the reasons that lead researchers to pursue
-%unsupervised
-%approaches to morphology; see e.g. \citet{BoStEnglishHindiTranslation2008}.
-Fortunately, we also had the opportunity to work with a CRF-based tagger from IIT
-Mumbai \citep{Kuhoo-Gupta:2006gd}.\footnote{Many thanks to Pushpak
-Bhattacharyya for making this tool available to us.}
-%dz 14.10.2009: následující URL možná už nefunguje?
-%\footnote{\url{http://www.cfilt.iitb.ac.in/tools/Hindi_POStagger.zip}}
-
-%The java application providing morphological analyses turned out to be a good
-%detector of dirty input data, because it crashed on various characters that
-%(mostly) did not belong to the input text. Most of these characters probably
-%appeared in the input text because of failed transcription as described in
-%\Sref{sec:tides}. However, since it crashes on occurrences of \% or \$, or
-%square brackets, which morphology uses for its own purpose, it can pose a
-%problem when these characters occur legitimately in the input text.
-
-%Once we cleaned-up and normalized the input data, morphology and POS tagger ran smoothly.
-
-\microsection{Textbook suffixes}
-\label{sec:textbook}
-Since the main morphological device of Hindi is alternating word suffixes,
-one can model suffixes instead of morphological tags.
-Our unsupervised approaches (see below) automatically acquire long but noisy
-lists of Hindi suffixes, some of which may bear grammatical meaning.
-As an alternative, we compiled a short list of 31 suffixes (including
-duplicates) for regular declension of nouns, verb formation etc.
-The cost of such a list was negligible: one of the authors (who does not speak
-Hindi) spent less then two hours with a Hindi textbook
-\citep{teach_yourself_hindi}, looking for grammar-describing sections.
-
-\subsection{Unsupervised Morphology} 
-We have tried two simple unsupervised methods:
-\begin{itemize}
-\item Automatic classes created by \texttt{mkcls}\footnote{\url{http://www.fjoch.com/mkcls.html}} tool contained in GIZA++ installation. 
-\item simple n-character long suffixes
-\end{itemize}
-and two more elaborate methods:
-%Affisix and Hindomor.
-
-\label{sec:aff}
-{\bf Affisix} is an open source tool for unsupervised recognition of affixes in a
-given language. It takes a large list of words and generates a list
-of possible suffixes or prefixes scored according to a selected method or a
-combination of methods. In this case, we used it to obtain the 100 top scoring
-suffixes
-using backward entropy and backward difference entropy methods (see
-\cite{affisixtsd2008} for details). The longest possible suffix seen in a word
-was then treated as the morphological tag of the word, leaving some words with a
-blank tag.
-
-\label{sec:hindomor}
-{\bf Hindomor} is another tool for unsupervised segmentation of words into
-morphemes. It was originally published in the context of information
-retrieval \citep{Zeman:2008}.
-
-The tool has been trained on the word types of the Hindi side of the Tides
-corpus. For every word the algorithm searches for positions where it can be cut
-in two parts: the stem and the suffix. Then it tries to filter the stem and
-suffix candidates so that real stems and suffixes remain. The core idea is that
-real stems occur with multiple suffixes and real suffixes occur with multiple
-stems.
-
-Given the lists of stems and suffixes obtained during training, we want to find
-the stem-suffix boundary in a word of the same language. Theoretically, we could
-use the learned stem-suffix combinations to require that both stem and suffix be
-known. However, this approach proved too restrictive, so we ended up in using
-just the list of suffixes. If a word ends in a string equal to a known suffix,
-the morpheme boundary is placed at the beginning of that substring.
 
@@
 
-The best BLEU score we achieved with Moses while making experiments with different
-factors for target side morphology was 12.22±0.78:
-trained on Tides and Daniel Pipes data with \texttt{hitbsuf}
-as the additional factor.
-
 \section{Hierarchical Phrase-Based Models}
@@
 
-One of the major differences between English and Hindi is the word order.
-Massive reordering must take place during translation.
-That is why we conducted several experiments with hierarchical
-translation models \citep{chiang:2007}, namely with Joshua \citep{joshua:2009},
-and compared the results with classical phrase-based models \citep[Moses, ][]{moses:2007}.
-
-Hierarchical phrase-based translation is based on synchronous context-free grammars (SCFG).
-Like phrase-based translation, pairs of corresponding source- and target-language phrases
-(sequences of tokens) are learnt from training data. The difference is that in hierarchical
-models, phrases may contain ``gaps'', represented by nonterminal symbols of the SCFG.
-If a source phrase $f$ contains a nonterminal $X_1$, then its translation $e$ also
-contains that nonterminal, and the decoder can replace the nonterminal by any phrase $f_1$
-and its translation $e_1$, respectively. An illustrative rule for English-to-Hindi translation is
-
-%Zatím neumíme hindštinu uvnitř matematiky, ale neškodilo by, kdybysme ji tam nějak dostali.
-%$$X \rightarrow \langle X_1 of X_2 , X_2 \hindi{का} X_1 \rangle$$
-
-\begin{center}
-$X \rightarrow \langle X_1\cdot of\cdot X_2\rangle , \langle X_2\cdot \mathsf{का} \cdot X_1 \rangle$
-\end{center}
-
-where the Hindi word \hindi{का} \textit{(kā)} is one of several grammatical forms of the Hindi equivalent
-of ``of'', and the subscripts on the nonterminals cause the two (noun) phrases around \textit{of}
-to be reordered around \hindi{का} in the translation.
-
-Hierarchical models have been known to produce better results than classical phrase-based
-models \citep{chiang:etal:2005}.
-
-In our experiments we used the Joshua implementation of the hierarchical phrase-based
-algorithms.\footnote{Many thanks to the team at Johns Hopkins University for creating
-Joshua and making it publicly available.} We set the maximum phrase length to 5,
-MERT worked with 300 best translation hypotheses per iteration, and the number of
-iterations was limited to 5.
-
-Unless stated otherwise, our experiments use a trigram language model trained on the
-target side of the Tides training data. In accord with \citet{BoStEnglishHindiTranslation2008},
-we found additional out-of-domain language models damaging to BLEU score.
-
-Symmetrical word alignments for grammar extraction were obtained using Giza++
-and the scripts accompanying Moses. Alignments were computed on first four
-(lowercased) characters of each training token, and the \textit{grow-diag-final-and}
-symmetrization heuristic was used.
-
-The results of the hierarchical models trained on various datasets,
-compared with classical phrase-based models are shown in \Tref{tab:joshua}.
-
-% - Popsat důležité varianty experimentů, které jsme zkoušeli:
-% -- Oficiální testování Ondrovým počítadlem (Ondra se zaručil, že na tokenizaci to nesahá):
-%    cp ~bojar/diplomka/granty/hindi/statmt/playground/exp.mert.SRCtides.train+en+lc.TGT+hi+for.c9ce7fbc.-fe.r0-0DEVtides.dev.Sbeam.20090806-1700/testbleu .
-%    gunzip -c /net/projects/hindi/data/augmented_corpora/tides.test/hi.gz > test.hi
-%    testbleu test.hi < /net/projects/hindi/experiments/joshua/josh11-tmtides-lmtides-lm3-preicon2-prenorm-pretokdz/eval/test.joshua.hi
-% -- baseline: trénování na tides (srovnat s Mosesem letos i loni)
-%    josh11-tmtides-lmtides-lm3-prebase = 11.28 (Ondra naměřil 11.10±0.79) (pozor, tenhle výsledek nemám s Ondrovým alignmentem lcstem4!)
-%    josh11-tmtides-lmtides-lm3-preicon2-prenorm-pretokdz = 12.27 (po normalizaci - kruci, měli bychom ukázat i nějaký výsledek před normalizací, jinak naše tvrzení, že jsme lepší než kdokoli loni, není zajímavé)
-%    testbleu: 12.27±0.83
-%    Moses: 1*LM 2*tides.train                                                     	11.46±0.72
-% -- trénování na tides a daniel pipes
-%    ac-tmtides+dpipes-lmtides-lm3 (allcstem4, ale tehdy jsem to ještě do názvu pokusu nepsal)
-%    = 12.58
-%    testbleu: 12.58±0.77
-%    Moses: 1*LM 2*tides.train 1*dani                                              	11.93±0.75
-% -- trénování na tides, daniel pipes a emille
-%    ac-allcstem4-tmtides+dpipes+omille-lmtides-lm3 = 11.32
-%    testbleu: 11.32±0.74
-%    Moses: 1*LM 2*tides.train 1*dani 1*emille-om                                  	10.06±0.72
-% -- trénování na tides, daniel pipes a slovník (zkusit ještě ten opravený?!!!)
-%    ac-allcstem4-tmtides+dpipes+dictfilt-lmtides-lm3 = 12.43 (to je starý neopravený a malý slovník)
-%    testbleu: 12.43±0.79
-%    Moses: 1*LM 2*tides.train 1*dani 1*dictfilt                                   	11.90±0.78
-% -- ke všem číselným výsledkům v této kapitole by to chtělo hned vedle vidět odpovídající výsledky Mosese.
-% PS: V sekci o Joshuovi by se mi libilo, kdybys to nejak propojil s lonskym reorderingek: ze vlastne podstata SCFG je svym zpusobem syntakticky transfer, tedy to, co jsme vloni hrube resili reorderingovymi pravidly.
+Joshua
 
@@
 
-\section{Results}
-\label{results}
-
-As far as automatic evaluation is concerned, the best result reported on in
-this paper is the 12.58 BLEU of Joshua trained on Tides and Daniel Pipes
-(\Tref{tab:joshua}). Moses was not able to outperform this score despite
-its ability to learn factored models. The best Moses score is 12.22 (morphology).
-%(in
-%tandem with the Mumbai tagger, see \Tref{tab:morph})
-
-The greatest mystery is the fact that adding Emille to the training data
-does not improve results with either system.
-% Joshua: tides+dpipes+omille: 0.1298 dev, 0.1132 test.
-% Joshua/Emille/Gaurav:        0.1285 dev, 0.1045 test.
-% Bez omilla (tides+dpipes):   0.1124 dev, 0.1258 test.
-Although we are not able to explain this in full, we made a few observations:
-%\begin{itemize}
-%  \item
-
-{\bf 1.}
-When asked to extract its SFCG, Joshua needs to see the source
-        (English) side of the data to be decoded, and extracts only the rules
-        relevant to the dataset. Thus we extract one grammar for the
-        development data (used for minimum error rate training) and another
-        for the test data. While the development grammar changes when Emille
-        is added, the test grammar remains the same. As if Emille was unable
-        to approximate the test data in any way.
-%  \item
-
-{\bf 2.}
-MERT optimizes 5 feature weights of Joshua: the language model
-        probability, the word penalty (preference for shorter translations),
-        and three translation features: $P(e|f)$, $P_{lex}(f|e)$ and
-        $P_{lex}(e|f)$. When Emille is involved, MERT always pushes the
-        non-lexical translation probability extraordinarily high, and causes
-        overfitting. While for other experiments we usually saw better BLEU
-        scores on test data than on development data, the opposite was the case
-        with Emille.
-%\end{itemize}
-
-\section{Related Research}
-
-% Tohle zamlcime, stejne jsme v tabulce.
-%\perscite{BoStEnglishHindiTranslation2008} tried to improve a statistical
-%phrase-based baseline MT
-%system by adding more data, standard lexicalized and a heuristical rule-based
-%reordering and an explicit modeling of (unsupervised) morphological coherence
-%on the target side. Neither of the approaches helped much.
-
-\perscite{pushpak:2009} improve Hindi morphology by transfering additional
-information from English parse tree and semantic roles, in addition to some
-hard-coded syntax-based reordering. However, they use an unspecified corpus
-making their results incomparable.
-
-% XXX reorderingy? no jen jednoho zastupce, a jako ze Joshua leccos resi sam
-
-%\subsection{Previous Work}
-\label{previous}
-
-The ICON 2008 NLP Tools Contest \citep{nlptoolscontest:2008:summary}
-included translation of Tides test data. Although our retokenized and cleaned
-data make a direct comparison impossible, our preliminary experiments on the
-unclean data are comparable. In \Tref{tab:previous} we compare four results
-presented at ICON 2008 with our hierarchical model trained on (unclean) Tides,
-with a Tides trigram language model.
-
-\begin{table}[ht]
-\begin{centering}
-\begin{tabular}{l|r}
-\textbf{System} & \textbf{BLEU} \\
-\hline
-Mumbai \citep{Damani:2008} & 8.53\\
-Kharagpur \citep{goswami:2008} & 9.76\\
-Prague \citep{BoStEnglishHindiTranslation2008} & 10.17\\
-Dublin \citep{srivastava:2008} & 10.49\\
-present Joshua & 11.10\\
-\end{tabular}
-\caption{Previous work compared to our hierarchical model (Joshua) on unclean data}
-\label{tab:previous}
-\end{centering}
-\end{table}
-
-\section{Human Evaluation}
-\label{human}
-
-The richer morphology and a high degree of reordering are known to render BLEU
-unreliable. While we still optimize for BLEU, our manual analysis reveals
-relatively low correlation. An extensive study similar to
-\cite{callisonburch-EtAl:2009:WMT} would be very valuable.
-
-We were able to conduct three small-scale manual evaluations. Our
-annotator was given the source English text and four or five Hindi translations in a
-randomized order of 100
-sentences (each time a different subset of the full test set). He assigned a
-mark to each of the four Hindi translations: giving no
-mark  (``0'') indicates a completely incomprehensible translation. A single star
-(``*'') denotes sentences with severe errors and hard to understand but still
-related to the input. Two stars (``**'') are assigned to more or less
-acceptable translations, possibly with many errors but understandable and
-conveying most of the original meaning.
-
-\Tref{tab:human} evaluates some basic configurations of Moses. For a comparison,
-we include the reference translations (REF) among the hypotheses (due to the
-randomization, the annotator cannot be entirely sure which of the hypotheses is
-the reference but often it can be guessed from the striking difference in
-quality). We see that even 6 reference translations were marked as inadequate
-and 11 as very bad. Because the test set is a part of the Tides corpus, these
-figures give a rough estimate of the overall corpus quality.
-
-The remaining figures in \Tref{tab:human} document that (while somewhat
-suspicious by itself), the Tides training data are most informative with respect
-to the (Tides) test set: a system trained completely out-of-domain on everything
-except Tides was able to deliver only 3 acceptable translations (OOD). The TIDP
-and WC10 systems compare the effect of adding more parallel data (TIDP for
-Tides+DanielPipes) vs. adding the unsupervised morphological factor (WC10, 10
-word classes). We observe that the difference indicated by human evaluation is
-rather convincing: nearly twice as many acceptable sentences in TIDP, but
-negligible in BLEU. This confirms our doubts about the utility of BLEU scores
-for languages with rich morphology and the neccessity to regularly run manual
-evaluations.
-
-\begin{table}[ht]
-\begin{centering}
-\begin{tabular}{l|r|r|r|r}
-{\bf System} & {\bf 0} & {\bf*} & {\bf**} & {\bf BLEU}\\
-\hline
-REF  &  6    &  11  &   83   &  -- \\
-OOD &   80  &   17   &  3   &   1.85±0.24\\
-TIDP &  26  &   44   &  30   &  11.93±0.75\\
-WC10 &  38   &  46   &  16   &  11.76±0.74\\
-\end{tabular}
-\caption{Manual evaluation of some basic Moses setups: training out of domain
-(OOD) and adding either more parallel data (TIDP) or a factor for morphological
-coherence (WC10) compared to the quality of the reference translation (REF).
-}
-\label{tab:human}
-\end{centering}
-\end{table}
-
-In \Tref{tab:human:mojo}, we evaluate the difference between phrase-based
-(Moses) and hierarchical (Joshua) translation model. Both systems are trained in
-identical conditions: both the translation and the language model are trained on
-Tides and DanielPipes. For Moses, we wanted to confirm the observation that more
-data are more important than ensuring morphological coherence on another 100
-manually judged sentences. Therefore, we report also Moses-DPipes+POStags, a
-setup trained on Tides only but including the supervised morphology of Mumbai
-tagger.
-
-We see that while the BLEU scores indicate the superiority of the hierarchical
-model over the phrase-based model with lexicalized reordering, the difference in
-manual judgments seems less convincing. Only 3
-sentences were ranked better. On the other hand, we confirm that even the
-supervised morphology is less important than a good parallel data source.
-
-\begin{table}[ht]
-\begin{small}
-\begin{centering}
-\begin{tabular}{l|r|r|r|r}
-{\bf System} & {\bf 0} & {\bf*} & {\bf**} & {\bf BLEU}\\
-\hline
-REF   &  6   &   10    &  84   &  -- \\ 
-Joshua  &  32   &  37   &  31   &   12.58±0.77\\
-Moses  &  35  &   35   &  30   &  11.93±0.75\\
-%Moses+DPipesR  &  33  &   42  &   25  &   11.80±0.73\\
-% Moses+DPipesR je reduced English, zatajuju
-Moses-DPipes+POStags  & 32  &   42   &  26   &  12.03±0.75\\
-\end{tabular}
-\caption{%
-Manual judgements of hierarchical (Joshua) vs. phrase-based (Moses) translation.
-}
-\label{tab:human:mojo}
-\end{centering}
-\end{small}
-\end{table}
-
-\Tref{tab:human:more} provides manual analysis of our baseline Moses setup
-(simple
-phrase-based translation, lexicalized reordering, language model trained on the
-full target side of the parallel data) trained on various subsets of
-our parallel data. We start with the combination of Tides and DanielPipes
-(TIDP), add Emille (EM), all other corpus sources (oth) and finally the
-dictionary Shabdanjali in two variants: full (DICTFull) and filtered to contain
-only Hindi words confirmed in a big monolingual corpus (DICTFilt). The BLEU
-scores indicate that Emille hurts the performance when tested on Tides test
-set. This surprising result was indeed confirmed by the manual analysis: only 12
-instead of 19 sentences were translated acceptably. Adding further data reduces
-the detrimental effect of Emille (not observed in BLEU scores) but the best
-performance is achieved by
-Tides+DanielPipes only.  Note that this final manual evaluation was based on a
-smaller dataset of 53 sentences only.
-
-\begin{table}[ht]
-\begin{small}
-\begin{centering}
-\begin{tabular}{l|r|r|r|r}
-{\bf System} & {\bf 0} & {\bf *} & {\bf**} & {\bf BLEU}\\
-\hline
-REF         &  0 &          8   &   45   & -- \\
-TIDP           &  20 &       14   &  19   &11.89±0.76\\
-TIDPEM        &   22 &          19   &   12   &9.61±0.75\\
-TIDPEMoth      &   17 &         25  &    11   & 10.97±0.79\\
-TIDPEMothDICTFilt     &    23 &  17  &    13  & 10.96±0.75\\
-TIDPEMothDICTFull    &    22 &   16  &    15  & 10.89±0.69 \\
-\end{tabular}
-\caption{The effect of additional (out-of-domain) parallel data in phrase-based
-translation.}
-\label{tab:human:more}
-\end{centering}
-\end{small}
-\end{table}
-
-\section{Conclusions and Future Work}
+\section{Conclusion}
 \label{sec:concl}
-%Do závěru bychom mohli plácnout něco jako ``We believe that most techniques
-%described in this paper are straightforwardly generalizable to other Indian SOV
-%languages.''
-
-We have tried several ways to improve state-of-the-art in English-Hindi SMT,
-however results are mixed:
-
-There are quite a few strange results: 
-\begin{itemize}
-\item POS tagging does not give better results than automatic word classes.
-Hindi textbook-based manual word classes were even better.
-\item Adding more training data to Tides either helps insignificantly, or (more
-often) hurts BLEU score. 
-\item BLEU may not correlate with human judgement. Our limited experiments show
-that adding training data may hurt BLEU but improve quality by human judgement.
-%\item It has been shown before by \citet{BoStEnglishHindiTranslation2008} that 
-%big monolingual Hindi data for LM didn't help either.
-\end{itemize}
-
-In our future work we want to further explore problems with existing datasets,
-the use of morphology, and the relation of output quality measured in terms of
-BLEU vs.
-human
-judgement. We also believe that there is room for improvement in the quality and
-amount of available parallel data.
 

### 2009-10-27 papers/2009-icon-hyderabad/paper.tex
Doplnen uvod, popsany vsechny 3 parsery i hlasovaci superparser.

@@
 \title{Maximum Spanning Malt: Hiring World's Leading Dependency Parsers to Plant Indian Trees%
+% Tohle nechat zakomentované, je to jen tahák, jak udělat acknowledgement grantu při nedostatku místa. Jinak ale mám momentálně na konci opravdovou sekci Acknowledgements.
 %\thanks{ \hspace{.6em}The research has been supported by the grant 
@@
 
-% Tady je posuzování taky slepé? Až odtajním autora, nezapomenout odtajnit i grant!
+% Tady je posuzování taky slepé? Až odtajním autora, nezapomenout odtajnit i acknowledgements!
 \author{%Daniel Zeman\\
@@
 
-Introduction goes here.
-\XXX Sections: Brief system description, Error analysis, Results, Conclusion
+Dependency parsing, i.e.
+sentence analysis that outputs tree of word-on-word dependencies (as opposed to constituent trees of context-free derivations), gained growing attention and popularity recently.
+There are data-driven dependency parsers that can be trained on syntactically annotated corpora (treebanks) and new, previously unseen material can be parsed very efficiently \citep{nivre:2009:ACLIJCNLP}.
+
+Most of the successful parsers employ discriminative learning techniques to sort out vast sets of potentially useful features observed in the input text.
+Thus, for every new training treebank, smart feature engineering is the key to getting the most out of the existing parsers, regardless how well they performed on other data sets and languages.
+Now that there are new treebanks available for two Indo-Aryan and one Dravidian language, we took three existing dependency parsers and explored the possibilities of tuning them for the new training data.
+Both parser configuration and data preprocessing are relevant approaches to the tuning.
+In addition, we used parser combination to further improve the results.
+
+Throughout the paper we focus mainly on the unlabeled attachment score. Although the parsers produce labeled dependencies, we do not optimize the system towards label accuracy.
+
+The rest of the paper is organized as follows: In \Sref{sec:system}, we describe the respective parsers and the combined parsing system.
+In \Sref{sec:experiments}, we report on the experiments we performed, discuss various results on the development set and analyze the errors.
+We conclude by summarizing the best configuration we were able to find, and future implications.
 
@@
 
-\XXX Acknowledgements: poděkovat Joakimovi a Ryanovi za zveřejnění parserů.
-\XXX other dependency parsers available for download include the isbn parser by titov
-\microsection{Tides}
-\label{sec:tides}
-We are happy
-to share our cleaning tools to make the experiments reproducible.\footnote{\url{https://wiki.ufal.ms.mff.cuni.cz/pub-company:icon2009}; accept the certificate.}
+DZ Parser builds a model of bigrams of words that occur together in a dependency; most of the time, words are identified by their part of speech tags and morphological features.
+The parser was originally developed for Czech but it can be re-trained for any other language.\footnote{Of course there are other dependency parsers that successfully participated in the CoNLL shared tasks and are available for download.
+One alternative worth mentioning is the ISBN Parser \citep{titov-henderson:2007:EMNLP-CoNLL2007} at \url{http://flake.cs.uiuc.edu/~titov/}.}
 
-\microsection{Daniel Pipes}
-\label{sec:dpipes}
-A journalist Daniel Pipes' website:\footnote{\url{http://www.danielpipes.org/}} limited-domain articles about the Middle East. 
-The articles are originally written in English, many of them are translated to
-up to 25 other languages, including Hindi (322 articles, 6,761 sentence
-pairs).
+\subsection{Voting Superparser}
+\label{sec:voting}
+The three parsers are combined using a simple weighted-voting approach similar to \citet{biblio:ZeZaImprovingParsing2005}, except that the output is guaranteed to be cycle-free.
+We start by evaluating every parser separately on the development data.
+The UAS of each parser is subsequently used as the weight of that parser's vote.
+Dependencies are parent-child relations, and for every node there are up to three candidates for its parent (if all three parsers disagree).
+Candidates get weighted votes -- e.g., if parsers with weights $w_1 = 0.8$ and $w_2 = 0.7$ agree on the candidate, the candidate gets 1.5 votes.
+Since we have only three parsers, in practice this means that the candidate of the best parser looses only if 1.
+the other two parsers agree on someone else, or 2.
+if attaching the child to this candidate would create a cycle.
 
-\subsection{Normalization}
+The tree is constructed from the root down.
+We repeatedly add nodes whose winning parent candidates are already in the tree.
+If none of the remaining nodes meet this condition, we have to break a cycle.
+We do so by removing the candidate (for parentship of any remaining node) with least weighted votes.
+Then we go on with adding nodes until all nodes are attached or there is another cycle to break.
+
+\section{Experiments}
+\label{sec:experiments}
 
@@
 
-\section{Hierarchical Phrase-Based Models}
-\label{joshua}
-
-Joshua
-
 \begin{table}[ht]
@@
 
+\section{Parser Unique Errors (Oracle Accuracy)}
+
+\section{How many times a candidate lost due to cycle prevention, and how many times this introduced an error?}
+
+\section{Learning Curve}
+
 \section{Conclusion}
@@
 
-On the other hand, we have shown that our hand-crafted word classes and some
+\XXX On the other hand, we have shown that our hand-crafted word classes and some
 additional data help Moses achieve significantly better results than reported
@@
 
-%\XXX -- někde zmínit, že "čištění" (ve smyslu Mosese, tedy dlouhých vět)
-%výsledky kazí, protože nám zmenšuje trénovací data. O to víc je vlastně
-%zvláštní, že přidání trénovacích dat typu Emille výsledky taky kazí.
+\section*{Acknowledgements}
+
+\XXX Acknowledgements: poděkovat Joakimovi a Ryanovi za zveřejnění parserů.
+The research has been supported by the grant 
+MSM0021620838 (Czech Ministry of Education).
 

### 2009-10-27 papers/2009-icon-hyderabad/paper.tex
Ve clanku uz je rada vysledku experimentu, ale jeste chybi vysledky s padem+vibhakti, jakoz i vsechny vysledky hlasovani.

@@
 In \Sref{sec:experiments}, we report on the experiments we performed, discuss various results on the development set and analyze the errors.
+In \Sref{sec:evaluation} we present the official results on the test data.
 We conclude by summarizing the best configuration we were able to find, and future implications.
@@
 
-blabla
+The final test data are blind, any error analysis is therefore impossible.
+That is why all scores given in this section were measured on the development data.
+All three treebanks follow the same annotation scheme and each of them is available in two flavors:
 
 \begin{compactitem}
-\item \hi{स्टैंडर्डज} \textit{(sṭaiṁḍarḍaja)} \translit{(sṭaiṁḍarḍaja)}
-\item \hi{स्टैंडर्डस} \textit{(sṭaiṁḍarḍasa)}
-\item \hi{स्टैंडर्ड्स} \textit{(sṭaiṁḍarḍsa)}
-\item \bn{সরিয়ে} Tohle je bengálsky.
-\item \te{వాడుతున్నాం} Tohle je telugsky.
+\item \textit{nomorph} variety contains word forms, chunk labels, dependency links and dependency labels
+\item \textit{morph} variety is augmented by automatically assigned lemmas, part of speech tags and values of morphological features (gender, number, person, case, postposition and tam -- tense+aspect+modality)
 \end{compactitem}
 
-\begin{table}[t]
-\begin{center}
-\small
-\begin{tabular}{l  l | ll}
-factor & BLEU &
-factor & BLEU\\
-\hline
-tag & 12.03±0.75 &	hitbsuf & 11.58±0.74\\
-wc50 & 11.97±0.73 &	hindomor2 & 11.55±0.74\\
-wc10 & 11.76±0.74 &	hindomor1 & 11.54±0.71\\
-lcsuf3 & 11.66±0.75 &	affddf & 11.50±0.7\\
-lcsuf1 & 11.63±0.72 &	affbdf & 11.33±0.72\\
-hindomor3 & 11.60±0.73 &	lcsuf2 & 11.14±0.74\\
-\end{tabular}
-\end{center}
-\caption{Target side morphology: Using different additional factors for second
-language model of MT system and its  effect on BLEU score. Trained on
-IIIT-TIDES only. 
-tag -- POS tags; wc\textit{n} -- \textit{n} word classes from mkcls; hitbsuf --
-word classes created by hand; lcsuf\textit{n} -- simple
-n-character suffixes; hindomor\textit{n};
-aff\textit{xxx} -- Affisix
-} 
-\label{tab:morph}
-\end{table}
+\Tref{tab:baseline} shows baseline results on the \textit{nomorph} data.
 
@@
 \begin{centering}
-\begin{tabular}{l|l|l}
-\textbf{Parallel data} & \textbf{Joshua} & \textbf{Moses} \\
+%\small
+\begin{tabular}{l|l|l|l}
+& \textbf{MST} & \textbf{Malt} & \textbf{DZ} \\
 \hline
-Tides & 12.27±0.83 & 11.46±0.72\\
-Tides+DP & \textbf{12.58±0.77} & 11.93±0.75\\
-Tides+DP+Emille & 11.32±0.74 & 10.06±0.72\\
-Tides+DP+Dict & 12.43±0.79 & 11.90±0.78\\
+hi & 80.32 & \textbf{81.84} & 62.00\\
+bn & 82.00 & \textbf{84.71} & 71.02\\
+te & 77.63 & \textbf{80.89} & 70.52\\
 \end{tabular}
-\caption{Results of Joshua compared with Moses}
-\label{tab:joshua}
+\caption{Baseline UAS of the three parsers on \textit{nomorph} development data. Language codes follow ISO 639: hi = Hindi, bn = Bangla, te = Telugu.}
+\label{tab:baseline}
 \end{centering}
@@
 
-\section{Parser Unique Errors (Oracle Accuracy)}
+\XXX what configuration of MST and Malt led to the baseline results?
 
-\section{How many times a candidate lost due to cycle prevention, and how many times this introduced an error?}
+Using the \textit{morph} data substantially deteriorates the accuracy.
+It could be caused by tagging errors (recall: morphological information has been disambiguated automatically) but data sparseness seems to be a more likely reason.
 
-\section{Learning Curve}
+\begin{table}[ht]
+\begin{centering}
+%\small
+\begin{tabular}{l|r|r|r|r|r|r}
+& \textbf{occ} & \textbf{frm} & \textbf{lem} & \textbf{cl} & \textbf{pos} & \textbf{feat} \\
+\hline
+hi & 13779 & 3973 & 3134 & 10 & \XXX & 714\\
+bn & 6449 & 2997 & 2336 & 14 & \XXX & 367\\
+te & 5494 & 2462 & 1403 & 12 & \XXX & 453\\
+\end{tabular}
+\caption{Size of the training corpora: occ -- word occurrences, frm -- distinct forms, lem -- lemmas, cl -- chunk labels, pos -- part of speech tags, feat -- feature value combinations.}
+\label{tab:corpus}
+\end{centering}
+\end{table}
+
+%\begin{table}[ht]
+%\begin{centering}
+%\small
+%\begin{tabular}{l|l|l|l}
+%& \textbf{MST} & \textbf{Malt} & \textbf{DZ} \\
+%\hline
+%hi & - & \textbf{-} & 44.32\\
+%bn & - & \textbf{-} & 50.31\\
+%te & - & \textbf{-} & 36.89\\
+%\end{tabular}
+%\caption{UAS on morph data (POS tag and features concatenated to one tag).}
+%\label{tab:baselinemorph}
+%\end{centering}
+%\end{table}
+
+\begin{table}[ht]
+\begin{centering}
+%\small
+\begin{tabular}{l|l|l|l}
+& \textbf{MST} & \textbf{Malt} & \textbf{DZ} \\
+\hline
+hi & - & \textbf{-} & 44.32\\
+bn & - & \textbf{-} & 50.31\\
+te & - & \textbf{-} & 36.89\\
+\end{tabular}
+\caption{UAS on morph data (POS tag and features concatenated to one tag).}
+\label{tab:poscasevib}
+\end{centering}
+\end{table}
+
+\XXX Otevřít soubor eval.txt. Mám v něm další výsledky, zejména u všech parserů natrénování na pos+pád+vibhakti.
+
+%\begin{compactitem}
+%\item \hi{स्टैंडर्डज} \textit{(sṭaiṁḍarḍaja)} \translit{(sṭaiṁḍarḍaja)}
+%\item \hi{स्टैंडर्डस} \textit{(sṭaiṁḍarḍasa)}
+%\item \hi{स्टैंडर्ड्स} \textit{(sṭaiṁḍarḍsa)}
+%\item \bn{সরিয়ে} Tohle je bengálsky.
+%\item \te{వాడుతున్నాం} Tohle je telugsky.
+%\end{compactitem}
+
+\subsection{Nonprojectivity}
+\label{sec:nonprojectivity}
+
+Nonprojectivity is a property of the dependency structure and the word order \citep{neproj} that makes parsing more difficult.
+All three parsers can produce nonprojective structures and all three treebanks are nonprojective.
+However, except for Hindi, the proportion of nonprojective dependencies is so small that one can hardly imagine that running the parsers in nonprojective mode would bring any improvement.
+A quick experiment with Malt Parser switched to the nonprojective Stack method revealed that it actually hurts the results even for Hindi.
+
+\begin{table}[ht]
+\begin{centering}
+%\small
+\begin{tabular}{l|ll}
+& \textbf{Edges} & \textbf{Sentences} \\
+\hline
+hi & 01.83 & 13.93\\
+bn & 00.96 & 05.49\\
+te & 00.45 & 01.31\\
+\end{tabular}
+\caption{Percentage of nonprojective dependencies, and of sentences containing at least one nonprojectivity.}
+\label{tab:nonprojectivity}
+\end{centering}
+\end{table}
+
+\subsection{Parser Unique Errors (Oracle Accuracy)}
+
+\subsection{How many times a candidate lost due to cycle prevention, and how many times this introduced an error?}
+
+\subsection{Learning Curve}
+
+\section{Official Evaluation}
+\label{sec:evaluation}
+
+Finally, we present the official evaluation of our voting superparser, as measured by the organizers on the test data.
+For this purpose, the parsing system has been retrained on both the training data and the development data.
+The results are shown in \Tref{tab:evaluation}.
+
+\begin{table}[ht]
+\begin{centering}
+\small
+\begin{tabular}{l|l|l|l}
+& \textbf{UAS} & \textbf{LAA} & \textbf{LAS} \\
+\hline
+hi & 88.58 (3:90.31) & 72.66 (4:76.38) & 68.60 (4:74.48)\\
+bn & 86.06 (4:90.32) & 71.28 (4:81.27) & 66.70 (5:79.81)\\
+te & 80.27 (4:86.28) & 54.20 (4:61.58) & 49.91 (4:60.55)\\
+\end{tabular}
+\caption{Official scores on the test data: unlabeled attachment score (UAS), label assignment accuracy (LAA) and labeled attachment score (LAS).
+The numbers in parentheses are the rank of our system and the score of the best system w.r.t.
+the given metric.}
+\label{tab:evaluation}
+\end{centering}
+\end{table}
 

### 2009-10-28 papers/2009-icon-hyderabad/paper.tex
Prvni verze clanku se blizi do cile, ale jeste musim pridat prinejmensim par cisel.

@@
 \newcommand{\te}[1]{{\tefont #1}}
-\newcommand{\translit}[1]{{\translitfont \textit{#1}}}
+\newcommand{\translit}[1]{{\translitfont \textit{(#1)}}}
 
@@
 \label{sec:mst}
-The Maximum Spanning Tree (MST) parser \citep{mst} views the sentence as an oriented complete graph with edges weighted by a feature scoring function.
-It finds for the graph a spanning tree that maximizes the weights of the edges.
+The Maximum Spanning Tree (MST) parser \citep{mst} views the sentence as an directed complete graph with edges weighted by a feature scoring function.
+It finds for the graph the spanning tree that maximizes the weights of the edges.
 A multi-class classification algorithm called MIRA is used to compute the scoring function.
@@
 The Malt Parser  \citep{malt} is a deterministic shift-reduce parser where input words can be either put to the stack or taken from the stack and combined to form a dependency.
-The decision, which operation to perform, is made by an oracle based on various features of the words in the input buffer and the stack.
+The decision which operation to perform is made by an oracle based on various features of the words in the input buffer and the stack.
 The default machine learning algorithm used to train the oracle is a sort of SVN (support vector machine) classifier \citep{svm}.
@@
 If none of the remaining nodes meet this condition, we have to break a cycle.
-We do so by removing the candidate (for parentship of any remaining node) with least weighted votes.
-Then we go on with adding nodes until all nodes are attached or there is another cycle to break.
+We do so by examining all unattached nodes.
+At each node we notice the votes of its current winning parent.
+Then we remove the least-scoring winner and go on with adding nodes until all nodes are attached or there is another cycle to break.
 
@@
 
-\Tref{tab:baseline} shows baseline results on the \textit{nomorph} data.
+\subsection{Morphology}
+\label{sec:morphology}
+
+\Tref{tab:baseline} shows baseline results on the \textit{nomorph} data. Both MST and Malt parsers were invoked in projective mode, Malt with the default Nivre arc-eager algorithm.
 
@@
 %\small
-\begin{tabular}{l|l|l|l}
-& \textbf{MST} & \textbf{Malt} & \textbf{DZ} \\
+\begin{tabular}{l|l|l|l|l}
+& \textbf{MST} & \textbf{Malt} & \textbf{DZ} & \textbf{Vote} \\
 \hline
-hi & 80.32 & \textbf{81.84} & 62.00\\
-bn & 82.00 & \textbf{84.71} & 71.02\\
-te & 77.63 & \textbf{80.89} & 70.52\\
+hi & 80.32 & 81.84 & 62.00 & \textbf{82.48}\\
+bn & 82.00 & \textbf{84.71} & 71.02 & 83.11\\
+te & 77.63 & \textbf{80.89} & 70.52 & 80.59\\
 \end{tabular}
-\caption{Baseline UAS of the three parsers on \textit{nomorph} development data. Language codes follow ISO 639: hi = Hindi, bn = Bangla, te = Telugu.}
+\caption{Baseline UAS of the four parsers on \textit{nomorph} development data. Language codes follow ISO 639: hi = Hindi, bn = Bangla, te = Telugu.}
 \label{tab:baseline}
@@
 
-\XXX what configuration of MST and Malt led to the baseline results?
-
-Using the \textit{morph} data substantially deteriorates the accuracy.
-It could be caused by tagging errors (recall: morphological information has been disambiguated automatically) but data sparseness seems to be a more likely reason.
+There are several ways how to use the additional information from the \textit{morph} data.
+The easiest way, exploitable by all three parsers, is to combine the chunk label, the POS tag and the features into one tag string.
+The results (not presented here) are very poor.
+Although there may be tagging errors, the most likely cause is data sparseness.
+In \Tref{tab:corpus} we illustrate this by showing the numbers of unique values in the various attributes of treebank words.
 
@@
 \hline
-hi & 13779 & 3973 & 3134 & 10 & \XXX & 714\\
-bn & 6449 & 2997 & 2336 & 14 & \XXX & 367\\
-te & 5494 & 2462 & 1403 & 12 & \XXX & 453\\
+hi & 13779 & 3973 & 3134 & 10 & 33 & 714\\
+bn & 6449 & 2997 & 2336 & 14 & 30 & 367\\
+te & 5494 & 2462 & 1403 & 12 & 31 & 453\\
 \end{tabular}
@@
 
-%\begin{table}[ht]
-%\begin{centering}
-%\small
-%\begin{tabular}{l|l|l|l}
-%& \textbf{MST} & \textbf{Malt} & \textbf{DZ} \\
-%\hline
-%hi & - & \textbf{-} & 44.32\\
-%bn & - & \textbf{-} & 50.31\\
-%te & - & \textbf{-} & 36.89\\
-%\end{tabular}
-%\caption{UAS on morph data (POS tag and features concatenated to one tag).}
-%\label{tab:baselinemorph}
-%\end{centering}
-%\end{table}
+To fight sparseness, we could either restrict the tag to selected information, or split the information into multiple features learnt separately, or both.
+We first restricted the tag string to selected information.
+Parsing on other treebanks showed that POS with case is especially useful \citep{dzparser}.
+The other feature we selected is called \textit{vibhakti}, which partially corresponds to case suffix and partially to postposition.\footnote{The Indian treebanks at hand are unusual in that nodes do not always map to words.
+They represent chunks, with function words such as postpositions hidden in node attributes.}
+
+\Tref{tab:poscasevib} presents the results of restricting the tag string to POS+case+vibhakti.
+This move especially improves the MST Parser, which now outperforms Malt on Hindi and Bangla.
+Malt improves on Hindi but drops behind on Bangla and Telugu.
+DZ Parser also improves on Hindi and deteriorates elsewhere but there is an interesting observation: even though its own scores are worse, the worsened output actually improves the voting results (provided the configurations of MST and Malt are fixed).
+So it seems that the newly introduced errors are less important (because taken care of by the more powerful parsers) while some difficult parts of the data are now covered better.
 
@@
 %\small
-\begin{tabular}{l|l|l|l}
-& \textbf{MST} & \textbf{Malt} & \textbf{DZ} \\
+\begin{tabular}{l|l|l|l|l}
+& \textbf{MST} & \textbf{Malt} & \textbf{DZ} & \textbf{Vote}\\
 \hline
-hi & - & \textbf{-} & 44.32\\
-bn & - & \textbf{-} & 50.31\\
-te & - & \textbf{-} & 36.89\\
+hi & 86.16 & 82.96 & 75.12 & \XXX 86.64\\
+bn & 85.70 & 80.02 & 54.38 & \XXX 78.79\\
+te & 79.85 & 78.37 & 45.78 & \XXX 78.81\\
 \end{tabular}
-\caption{UAS on morph data (POS tag and features concatenated to one tag).}
+\caption{UAS on refined \textit{morph} data (POS tag, case and vibhakti concatenated). \XXX voting se pocitalo, kdyz jeste MST pracoval jen s nomorph, takze to musime prepocitat}
 \label{tab:poscasevib}
 \end{centering}
+% Když se ke značce POS přilepí pád, ale ne vibhakti, a s tím se přetrénuje Malt parser (ostatní parsery jsme nechali beze změny, tj.
+MST byl natrénovaný na nomorph a DZ naopak i na vibhakti), tak Malt sám má na všech třech jazycích pořád stejné výsledky (jako v této tabulce), ale výsledky Vote se na hindštině zhorší a na zbývajících dvou jazycích zlepší: hi=85.36, bn=83.35, te=80.89.
 \end{table}
 
-\XXX Otevřít soubor eval.txt. Mám v něm další výsledky, zejména u všech parserů natrénování na pos+pád+vibhakti.
+Finally, ve returned to \textit{nomorph} with Malt on Bangla and Telugu, where the POS+case+vibhakti tags did not help.
+The results are given in \Tref{tab:posmix2}.
+This is also the configuration we used to parse the test set for the official evaluation round 1.
+
+\begin{table}[ht]
+\begin{centering}
+%\small
+\begin{tabular}{l|l|l|l|l}
+& \textbf{MST} & \textbf{Malt} & \textbf{DZ} & \textbf{Vote}\\
+\hline
+hi & 86.16 & 85.84 & 75.12 & \textbf{87.12}\\
+bn & 85.70 & 84.71 & 54.38 & \textbf{86.19}\\
+te & 79.85 & 80.89 & 45.78 & \textbf{82.37}\\
+\end{tabular}
+\caption{UAS on mixed data: MST and DZ use POS+case+vibhakti for all languages, Malt uses that for Hindi and POS only elsewhere.
+MST is now the best parser for hi and bn, which changes the voting weights.}
+\label{tab:posmix2}
+\end{centering}
+\end{table}
 
@@
 However, except for Hindi, the proportion of nonprojective dependencies is so small that one can hardly imagine that running the parsers in nonprojective mode would bring any improvement.
-A quick experiment with Malt Parser switched to the nonprojective Stack method revealed that it actually hurts the results even for Hindi.
+A quick experiment with Malt Parser switched to the nonprojective Stack Eager algorithm revealed that it actually hurts the results even for Hindi.
 
@@
 
-\subsection{Parser Unique Errors (Oracle Accuracy)}
+\subsection{Error Patterns}
+\label{sec:errors}
 
-\subsection{How many times a candidate lost due to cycle prevention, and how many times this introduced an error?}
+The accuracy of the dependencies is relatively high and it is difficult to trace repetitive error patterns.
+In Hindi, many wrong attachments seem to be long-distance, and verbs, conjunctions, root and NULL nodes are frequently involved.
+Frequent words should perhaps be available to the parsers as parts of tag strings: for instance, Hindi \hi{कि} \translit{ki} ``that'' or \hi{तो} \translit{to} are wrongly attached because the parser only sees the general CC tag.
+On a similar note, problems with coordination, also observed e.g.
+by \citet{dzparser}, occur here, too: \hi{भाई और भाभी} \translit{bhāī aura bhābhī} ``brother and his wife'' is correctly recognized as coordination rooted by the conjunction \hi{और}, however, the conjunction node lacks the information about its noun children and fails to attach as the subject of the verb.
 
-\subsection{Learning Curve}
+The tag string should contain both the chunk label and the POS tag.
+So far we wrongly assumed that POS always determins the chunk label.
+It is often so but not always, as exemplified in the Bangla chunk sequence \bn{তবে সুদীপ ওকে একদিন আড়ালে ডেকে বলেছিল কৌতূহল দেখালে তুমি উঁচুতে উঠতে অনিমেষ} \translit{tabe sudīpa oke ekadina āṛāle ḍeke balechila kautūhala dekhāle tumi um̃cute uṭhate animeša}.
+The words \bn{ডেকে} and \bn{দেখালে} are tagged VGNF|VM while \bn{বলেছিল} and \bn{উঠতে} are VGF|VM.
+The parser gets them wrong and it could be caused by it seeing only VM in the tag.
+
+In Telugu, extraordinal number of sentences follow the SOV order so strongly that the last node (verb) is almost always attached to the root and most other nodes are attached directly to the last node.
+An example chunk sequence where this rule would lead to 100~\% accuracy follows: \te{రాష్ట్రంలొ రంగారెడ్డి మెదక్ నిజామాబాద్ జిల్లాలలొ పంటను గొప్పొ పండిస్తున్నారు} \translit{rāšṭraṁlo raṁgāreḍḍi medak nijāmābād jillālalo paṁṭanu goppo paṁḍistunnāru}.
+In the light of such examples it seems reasonable to provide the parsers with an additional feature telling whether a particular dependency observes the ``naïve Telugu'' structure.
+Note however that this will not help with the other two languages.
+While 73.75~\% of Telugu dependncies follow this rule, it is only 39.52~\% in Bangla and 35.71~\% in Hindi.
+
+\subsection{\XXX To Do}
+\begin{compactitem}
+\item \XXX Parser unique errors (oracle accuracy)
+\item \XXX How many times a candidate lost due to cycle prevention, and how many times this introduced an error?
+\item \XXX Learning curve
+\item \XXX Spočítat, kolik je ve kterém jazyce uzlů NULL.
+\item \XXX znacka chunku, priklad z bengalstiny
+\item \XXX casta slova, priklad z hindstiny
+\item \XXX koordinace, priklad z hindstiny
+\end{compactitem}
 
@@
 
+\section{Related and Future Work}
+\label{sec:related}
+
+There is a large body of work on parser combination.
+A summary can be found in \citet{nivre-mcdonald:2008:ACLMain}, whose approach is also related to ours w.r.t.
+the selection of parsers.
+However, their feature-based integration of MST and Malt parsers is much more sophisticated than our lightweight voting.
+Further improvement of accuracy can be expected if MST-Malt integration is applied to the Indian treebanks.
+
+Future work, at the time of writing, includes experiments that can be run before the evaluation round 2, and thus their results will appear in the final version of this paper.
+We definitely intend to test other algorithms of the Malt parser, as well as to reconfigure its feature pool and let it work with POS, case and vibhakti separately.
+
+Labeling of the dependencies is another problem that deserves more attention.
+We have concentrated on the unlabeled attachment score so far and for the sake of the official evaluation, we simply pushed the MST labels through.
+A separate postprocessing classifier would probably produce better results.
+
 \section{Conclusion}
@@
 
-\XXX On the other hand, we have shown that our hand-crafted word classes and some
-additional data help Moses achieve significantly better results than reported
-previously. Hierarchical decoder Joshua can capture word order even better than
-Moses. Its results are always slightly better. And as far as we know, our
-current results are the best that have been reported on this dataset.
+We have described our system of voting parsers, as applied to the ICON 2009 NLP Tools Contest task.
+We showed that case and vibhakti are important features at least for parsing Hindi while their usability in Bangla and Telugu is limited by data sparseness.
+We also discussed several error patterns that could lead to further improvements of the parsing system in future.
 
@@
 
-\XXX Acknowledgements: poděkovat Joakimovi a Ryanovi za zveřejnění parserů.
-The research has been supported by the grant 
-MSM0021620838 (Czech Ministry of Education).
+We are enormously grateful to the developers of MST and Malt parsers for making their software available to the research community.
+
+% Tohle nezapomenout odkomentovat, až budeme posílat finální verzi!
+%The research has been supported by the grant 
+%MSM0021620838 (Czech Ministry of Education).
 

### 2009-10-30 papers/2009-icon-hyderabad/paper.tex
Clanek uz se asi muze odeslat Samarovi.

@@
 
-% Tady je posuzování taky slepé? Až odtajním autora, nezapomenout odtajnit i acknowledgements!
-\author{%Daniel Zeman\\
-%Univerzita Karlova v Praze, Ústav formální a aplikované lingvistiky\\
-%Malostranské náměstí 25, CZ-11800, Praha, Czechia\\ 
-%\texttt{zeman@ufal.mff.cuni.cz}
+\author{Daniel Zeman\\
+Univerzita Karlova v Praze, Ústav formální a aplikované lingvistiky\\
+Malostranské náměstí 25, CZ-11800, Praha, Czechia\\ 
+\texttt{zeman@ufal.mff.cuni.cz}
 }
@@
 \hline
-hi & 86.16 & 82.96 & 75.12 & \XXX 86.64\\
-bn & 85.70 & 80.02 & 54.38 & \XXX 78.79\\
-te & 79.85 & 78.37 & 45.78 & \XXX 78.81\\
+hi & 86.16 & 85.84 & 75.12 & \textbf{87.12}\\
+bn & 85.70 & 77.31 & 54.38 & \textbf{85.82}\\
+te & \textbf{79.85} & 77.78 & 45.78 & 79.70\\
 \end{tabular}
-\caption{UAS on refined \textit{morph} data (POS tag, case and vibhakti concatenated). \XXX voting se pocitalo, kdyz jeste MST pracoval jen s nomorph, takze to musime prepocitat}
+\caption{UAS on refined \textit{morph} data (POS tag, case and vibhakti concatenated).}
 \label{tab:poscasevib}
@@
 \end{table}
-
-%\begin{compactitem}
-%\item \hi{स्टैंडर्डज} \textit{(sṭaiṁḍarḍaja)} \translit{(sṭaiṁḍarḍaja)}
-%\item \hi{स्टैंडर्डस} \textit{(sṭaiṁḍarḍasa)}
-%\item \hi{स्टैंडर्ड्स} \textit{(sṭaiṁḍarḍsa)}
-%\item \bn{সরিয়ে} Tohle je bengálsky.
-%\item \te{వాడుతున్నాం} Tohle je telugsky.
-%\end{compactitem}
 
@@
 
-\subsection{\XXX To Do}
-\begin{compactitem}
-\item \XXX Parser unique errors (oracle accuracy)
-\item \XXX How many times a candidate lost due to cycle prevention, and how many times this introduced an error?
-\item \XXX Learning curve
-\item \XXX Spočítat, kolik je ve kterém jazyce uzlů NULL.
-\item \XXX znacka chunku, priklad z bengalstiny
-\item \XXX casta slova, priklad z hindstiny
-\item \XXX koordinace, priklad z hindstiny
-\end{compactitem}
+\subsection{Voting Potential}
+\label{sec:votingpotential}
+
+In order to see how much can be potentially gained from parser combination, we summarized the attachments that at least one of the parsers got correct.
+This \textit{oracle accuracy} gives an upper limit for the real scores we can achieve.
+It corresponds to the case that for every word, an oracle correctly tells which parser to ask about the word's parent.
+\Tref{tab:oracle} presents the oracle accuracies together with percentage of unique correct attachments that only one parser delivered.
+These figures give some idea of how much similar are the errors of the respective parsers to each other.
+Malt parser has the most unique know-how in all three languages, which could be explained by its focus on local features.
+Both MST and DZ can reach for global, sentence-wide relations.
+Note however, that the development data set is small and the percentages correspond to 42 (Malt/Bangla) or less words.
+
+\begin{table}[ht]
+\begin{centering}
+%\small
+\begin{tabular}{l|l|l|l|l}
+& \textbf{Oracle} & \textbf{UqMST} & \textbf{UqMalt} & \textbf{UqDZ} \\
+\hline
+hi & 93.92 & 2.96 & \textbf{3.12} & 1.84\\
+bn & 94.20 & 4.32 & \textbf{5.18} & 1.97\\
+te & 88.00 & 2.37 & \textbf{5.48} & 2.07\\
+\end{tabular}
+\caption{Oracle accuracy for the three languages, and unique correct attachments (\%) proposed by a single parser.}
+\label{tab:oracle}
+\end{centering}
+\end{table}
+
+%\subsection{\XXX To Do}
+%\begin{compactitem}
+%\item \XXX How many times a candidate lost due to cycle prevention, and how many times this introduced an error?
+%\item \XXX Learning curve
+%\item \XXX Spočítat, kolik je ve kterém jazyce uzlů NULL.
+%\item \XXX casta slova, priklad z hindstiny
+%\item \XXX Změřit výkon kombinovaného parseru. K tomu je potřeba rozchodit skript zmeritvykon.pl, nechtěl mi fungovat.
+%\end{compactitem}
 
@@
 We are enormously grateful to the developers of MST and Malt parsers for making their software available to the research community.
-
-% Tohle nezapomenout odkomentovat, až budeme posílat finální verzi!
-%The research has been supported by the grant 
-%MSM0021620838 (Czech Ministry of Education).
+The research has been supported by the grant 
+MSM0021620838 (Czech Ministry of Education).
 

### 2009-10-30 papers/2009-icon-hyderabad/paper.tex
Tak v teto podobe to odeslu Samarovi k posouzeni.

@@
 \label{sec:mst}
-The Maximum Spanning Tree (MST) parser \citep{mst} views the sentence as an directed complete graph with edges weighted by a feature scoring function.
+The Maximum Spanning Tree (MST) parser \citep{mst} views the sentence as a directed complete graph with edges weighted by a feature scoring function.
 It finds for the graph the spanning tree that maximizes the weights of the edges.
@@
 We do so by examining all unattached nodes.
-At each node we notice the votes of its current winning parent.
+At each node we note the votes of its current winning parent.
 Then we remove the least-scoring winner and go on with adding nodes until all nodes are attached or there is another cycle to break.
@@
 \end{tabular}
-\caption{UAS on mixed data: MST and DZ use POS+case+vibhakti for all languages, Malt uses that for Hindi and POS only elsewhere.
-MST is now the best parser for hi and bn, which changes the voting weights.}
+\caption{UAS on mixed data: MST and DZ use POS+case+vibhakti for all languages, Malt uses that for Hindi only, elsewhere it uses just POS.}
 \label{tab:posmix2}
@@
 The tag string should contain both the chunk label and the POS tag.
-So far we wrongly assumed that POS always determins the chunk label.
+So far we wrongly assumed that POS always determines the chunk label.
 It is often so but not always, as exemplified in the Bangla chunk sequence \bn{তবে সুদীপ ওকে একদিন আড়ালে ডেকে বলেছিল কৌতূহল দেখালে তুমি উঁচুতে উঠতে অনিমেষ} \translit{tabe sudīpa oke ekadina āṛāle ḍeke balechila kautūhala dekhāle tumi um̃cute uṭhate animeša}.
@@
 We showed that case and vibhakti are important features at least for parsing Hindi while their usability in Bangla and Telugu is limited by data sparseness.
+Providing these features to MST and DZ in all languages, and to Malt in Hindi only yielded the best combined parser.
 We also discussed several error patterns that could lead to further improvements of the parsing system in future.
